### 2024-02-12 outputs/paper/paper.Rmd
e

---
title: "Declining Birth rates in the US: An Analysis of Potential Factors"
subtitle: "Reproduction of 'The Puzzle of Falling US Birth Rates since the Great Recession' (Kearney, Levine & Pardue, 2022)"
author: Heng Ma
thanks: "Code and data are available at: https://github.com/Brian031205/paper_2.git. A replication of various aspects in this paper are available at: ______________    "
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Over the past 10 years, US birth rates have been dropping steadily to an all time low of 55.8 per 1,000 women in 2020. Kearney, Levine and Pardue (2022) analyzed this phenomenon by exploring demographic, economic, social and policy factors to little enlightenment of the cause of the decline. However, they did discover a correlation between the decline and the birth cohorts of mothers, and offer the conjecture of shifting priorities as the reasoning. We replicate the results of this study with respect to the demographic and cohort effects, and discuss alternative interpretations of the data considering the impact of the feminist movement and the evolution of technology and media. The results were consistent with the initial study."
output:
  bookdown::pdf_document2
toc: TRUE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = "h", out.extra = "")

# Load libraries
if(!require(Rcpp)){install.packages('Rcpp', dependencies = TRUE)}
#if(!require(jmv)){install.packages('jmv', repos=c('https://repo.jamovi.org', 'https://cran.r-project.org'))}
if(!require(jmv)){install.packages('jmv', dependencies = TRUE)}
if(!require(rstudioapi)){install.packages('rstudioapi', dependencies = TRUE)}
if(!require(effsize)){install.packages('effsize', dependencies = TRUE)}
if(!require(Hmisc)){install.packages('effsize', dependencies = TRUE)}
if(!require(psych)){install.packages('psych', dependencies = TRUE)}
if(!require(ggplot2)){install.packages('ggplot2', dependencies = TRUE)}
if(!require(dplyr)){install.packages('dplyr', dependencies = TRUE)}
if(!require(MBESS)){install.packages('MBESS', dependencies = TRUE)}

library(MBESS)
library(psych)
library(ggplot2)
library(dplyr) 
library(Hmisc)
library(effsize)
library(rstudioapi)
library(jmv)
library(Rcpp)

# Read in data

```

# Introduction

<!-- Background / Broader Context -->
Birth rate is the number of live births per thousand in the population by year [@dict]. Birth rate is important since it is an indicator of population growth and determines the age structure of populations, which has various implications for the economy and society [@CLELAND2008364]. When the birth rate is too high or low, it affects all subgroups of the population in different ways, for example, a large number of new births will be a burden on the adult population to support them, and when this group ages to be elderly they will be a burden on governments to support them [@CLELAND2008364].

There are many possible factors that can affect birth rate. Recessions have been proposed to result in a drop in birth rates, specifically the Great Recession of 2007 [@pew; @focus; @econ]. @becker proposes an economic framework for fertility, where the "demand for children" is driven by the satisfaction that children brings to people weighted against factors such as preferences (e.g. religion, race, age), income, time, expenditure and means to support each child and other considerations (e.g. contraception, marriage). @og base their analysis of the declining US birth rates on these factors.

These analyses however, did not provide sufficient evidence to explain the decline in birth rates, with the authors stating "For any factor to have explained much of that decline, it would have had to change dramatically around the same time." [@og]. @og then went on to find a correlation between the recent birth cohorts of mothers and the substantial decline. In our reproduction of their paper, we build on their 'shifting priorities' explanation of the cohort effects by considering the other notable events that potentially may have influenced the nature of parenting of these cohorts, namely the Third wave of Feminism and the second half of the Digital Revolution.

We replicate the paper by @og with a focus on the following research questions:

* What is the trend in US birth rates over the period 1980-2020?
* How does the trend in birth rates vary with demographic factors such as age, race, education and marital status?
* How do the birth rates vary for different cohorts of mothers by their birth year?

While the original paper used Stata [@citeStata] for data processing and analysis in its replication package, we use R [@citeR] for all data wrangling and analysis and R packages tidyverse [@tidy], ggthemes [@ggthemes], ggprism [@ggprism] and patchwork [@patchwork] to produce the figures, kableExtra [@kableExtra] to produce the tables and haven [@haven] to read the dta files.

# Data

## Data Source and Methodology

The data on birth rates per 1,000 women ages 15-44 across all races and demographic population subgroups are obtained from the National Vital Statistics Reports for the years 2015, 2019 and 2020 [@2015birth; @2019birth; @2020birth], which they collected from birth certificates registered in all  US states and the District of Colombia. Aggregated data on the number of births for 6 different 5-year cohorts of mothers by their age and birth year is provided by @og, calculated using public birth microdata across the period 1980-1989 from the NBER Natality Database and NCHS microdata from the period 1990-2019 [@nchsR; @nber]. The birth cohorts span the years 1968-1997. The NCHS microdata is restricted use and requires an application to be obtained. We simply used the data that was supplied in the replication packageby @og.

Single-age population counts, among all races from 1969-2019 and by race and Hispanic origin from 1990-2019, is obtained from the CDC SEER database [@nci].

## Attributes

The birth data consists of variables for the year, and birth rates for each population subgroup. For example, for race there are 3 variables: birth rate for whites, birth rate for blacks and birth rate for hispanics. There is birth rate data for 4 population subgroups - age (5-year groupings), race and ethnicity, marital status and mother's level of education. We show a glimpse of the data with variables for race and marital status in table \@ref(tab:birthdata).

```{r birthdata, echo=FALSE}
# Makes table of Ten rows of the birth rate data for years 1980-2020
birth_data %>%
  select(year, brate_all, brate_whitenh, brate_blacknh, brate_hisp, brate_unmarried, brate_married) %>%
  filter(year > 1989) %>%
  slice(1:10) %>%
  kable(
    caption = "Ten rows of the birth rate data for years 1980-2020",
    col.names = c("Year", "All races", "White, non-Hispanic", "Black, non-Hispanic", "Hispanic", "Unmarried", "Married"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  )
```

For the aggregated cohort birth data, the variables are mother's age, birth cohort, number of births and cumulative births. The population data provided by @og gives the population count for women by age, year and state. We group the population count by cohort and mother's age and join this to the cohort birth data, with the first 10 rows of this data shown in table \@ref(tab:birthcohort). We also added a column with the birth years for each cohort for easier interpretation.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Process population data to aggregate by cohort
fig5_pop_data1 <- pop_cohort_data %>%
  select(1:32) %>%
  select(-c(stname)) %>%
  group_by(year) %>%
  summarise(across(everything(),sum)) # https://stackoverflow.com/questions/1660124/how-to-sum-a-variable-by-group

fig5_pop_data1 <- fig5_pop_data1 %>%
  pivot_longer(-c(year), names_to = "mage", values_to = "pop")

fig5_pop_data1 <- fig5_pop_data1 %>%
  mutate(mage = as.numeric(str_replace(mage, "fem", ""))) %>%
  mutate(cohort = year - mage) %>%
  mutate(cohort2 = case_when(cohort >= 1968 & cohort <=1972 ~ 1,
                             cohort >= 1973 & cohort <=1977 ~ 2,
                             cohort >= 1978 & cohort <=1982 ~ 3,
                             cohort >= 1983 & cohort <=1987 ~ 4,
                             cohort >= 1988 & cohort <=1992 ~ 5,
                             cohort >= 1993 & cohort <=1997 ~ 6))

fig5_pop_data1 <- na.omit(fig5_pop_data1)
fig5_pop_data1 <- fig5_pop_data1 %>%
  group_by(cohort2, mage) %>%
  summarise(pop = sum(pop))

# Join dataframes for population and birth data
birth_cohort_data <- full_join(birth_cohort_data, fig5_pop_data1, by=c("mage", "cohort2")) %>%
  mutate(cohort_years = case_when(cohort2 == 1 ~ "1968-1972",
                                  cohort2 == 2 ~ "1973-1977",
                                  cohort2 == 3 ~ "1978-1982",
                                  cohort2 == 4 ~ "1983-1987",
                                  cohort2 == 5 ~ "1988-1992",
                                  cohort2 == 6 ~ "1993-1997"))

```

```{r birthcohort, echo=FALSE}
# Make table of First ten rows of the birth cohort data for 6 birth cohorts
birth_cohort_data %>%
  select(2,6,1,3,4,5) %>%
  slice(1:10) %>%
  kable(
    caption = "First ten rows of the birth cohort data for 6 birth cohorts",
    col.names = c("Cohort #", "Birth Years", "Mother's age", "Number of births", "Cumulative births", "Population Count"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  )
```

# Results

## Overall evaluation in regre depending hitchhiker scenario

```{r table1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Rates for Regret, Social norms and Negative affect"}

Orginal_table <- read.csv("/cloud/project/inputs/data/Ex1.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
print(Orginal_table)
```

Figure \@ref(fig:fig1) is a replication of Figure 1 in the original paper, In this section, participants evaluated a classic scenario known as the hitchhiker scenario, to assess the impact of past behavior on the perception of regret following negative outcomes. 92% of the 315 participants believed that Mr. Jones, who almost never gives rides to others, felt a stronger sense of regret. The majority of participants believed that Mr. Jones, who almost never picks up hitchhikers, exhibits behavior that is more common in society (95% out of 326 respondents), whereas Mr. Smith, who frequently gives rides to hitchhikers, is more likely to be criticized by society (91% out of 310 respondents). For the assessment of negative affect, 93% of people believed that Mr. Jones experienced more negative affect compared to Mr. Smith. Overall, these two experiments provided very close replications, similar to the the original study.

```{r fig1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Rates for Regret, Social norms and Negative affect"}
data1 <- read.csv("/cloud/project/inputs/data/Ex1.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
str(data)

# Figure 1 replication

library(ggplot2)
library(reshape2)
library(ggplot2)
library(readr)
library(dplyr)

data <- data.frame(
  Category = c('Regret', 'Social norms (injunctive)', 'Social norms (descriptive)', 'Negative affect'),
  Smith_Routine = c(7.9, 4.7, 90.6, 7.3),
  Jones_Exception = c(92.1, 95.3, 9.4, 92.7)
)

long_data <- melt(data, id.vars = 'Category', variable.name = 'Group', value.name = 'Percentage')

ggplot(long_data, aes(x = Category, y = Percentage, fill = Group)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  scale_fill_grey(start = 0.8, end = 0.2) + # Adjust the grey scale to match the example
  theme_minimal() +
  labs(
    fill = "", 
    x = "",    
    y = ""     
  ) +
  theme(
    legend.position = "bottom", 
    axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels if necessary
    plot.title = element_text(hjust = 0.5) # Center the plot title if you add one
  )
print(ggplot)
```

## Demographic Factors

Figure \@ref(fig:fig2) is a replication of Figure 2 in the original paper, and shows the Proportions for perceived regret and luck.This graph compares the relationship between regret and luck, with 81% of 277 respondents believing that Mr. White, who had just changed to a new route home and experienced an accident, would feel more regretful.
```{r table1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Rates for Regret, Social norms and Negative affect"}

table2 <- read.csv("/cloud/project/inputs/data/table2.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
print(table2)
```
 As for luck, Mr. White (exception) was considered less lucky than Mr. Adams (routine), with 114 participants (33%) viewing Mr. Adams as less lucky, whereas 228 (66%) perceived Mr. White to be less lucky.

```{r fig2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Trends in Birth Rates by Population Subgroup"}

library(ggplot2)
library(reshape2)

df <- read.csv("/cloud/project/inputs/data/table2.csv", header = TRUE, stringsAsFactors = FALSE)

# Clean up the data by removing unnecessary 'Counts' columns and renaming the remaining
df_clean <- df[, -c(2, 4)]
names(df_clean) <- c("Category", "Regret", "Luck")

# Melt the data frame to have 'Category', 'Variable', and 'Value' columns
df_melted <- melt(df_clean, id.vars = "Category", variable.name = "Outcome", value.name = "Percentage")

# Plot with ggplot2
p <- ggplot(df_melted, aes(x = Outcome, y = Percentage, fill = Category)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
  scale_fill_manual(values=c("Routine Adams" = "black", "Exception White" = "grey")) +
  labs(x = "", y = "Percentage", fill = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")

# Print the plot
print(p)
```

```{r proportions, echo=FALSE, warning=FALSE, message=FALSE}
this.dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(this.dir)

# load our dataset 
data <- read.csv("/cloud/project/inputs/data/osf-past-normality-regret-replication-exp1-data.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
str(data)

data$Sc1_regret<- factor(data$Sc1_regret)
data$sc1_socnorms1<- factor(data$sc1_socnorms1)
data$sc1_socnorms2<- factor(data$sc1_socnorms2)
data$sc1_combinednorms<- factor(data$sc1_combinednorms)

# Let's label the values better, so it's easier to understand the output.
data$Sc1_regret <- ordered(data$Sc1_regret, levels = c(1,2), labels = c("Exception Jones", "Routine Smith"))
data$sc1_socnorms1 <- ordered(data$sc1_socnorms1, levels = c(1,2), labels = c("Exception Jones", "Routine Smith"))
data$sc1_socnorms2 <- ordered(data$sc1_socnorms2, levels = c(1,2), labels = c("Exception Jones", "Routine Smith"))
data$sc1_combinednorms <- ordered(data$sc1_combinednorms, levels = c(1,2), labels = c("Exception Jones", "Routine Smith"))

data$Sc3condition1 <- 0
data$compensationagg1 <- 0
data$regretagg1 <- 0
for (i in 1:nrow(data)){
  if (!is.na(data$Sc3_C1_text[i])){
    data$Sc3condition1[i] <- 1
    data$compensationagg1[i] <- data$sc3_c1_compensation[i]
    data$regretagg1[i] <- data$sc3_c1_regret[i]
  }
  else if (!is.na(data$Sc3_C2_text[i])){
    data$Sc3condition1[i] <- 2
    data$compensationagg1[i] <- data$sc3_c2_compensation[i]
    data$regretagg1[i] <- data$sc3_c2_regret[i]
  }
  else if (!is.na(data$Sc3_C3_text[i])){
    data$Sc3condition1[i] <- 2
    data$compensationagg1[i] <- data$sc3_c3_compensation[i]
    data$regretagg1[i] <- data$sc3_c3_regret[i]
  }
  else {
    data$Sc3condition1[i] <- NA
    data$compensationagg1[i] <- NA
    data$regretagg1[i] <- NA
  }
}

#value labels
data$Sc3conditionl1<-factor(data$Sc3condition1,levels = c(1,2), labels=c("Routine", "Exception"))

# let's have a look at this
table(data$Sc3conditionl1)

#Adjust Copensation-Scale to original paper (Values from 0 to 10, instead of 1 to 11)
data$compensationaggrecoded1 = data$compensationagg1-1

#Adjust Regret Scale (Values 1 to 5, indead of 0 to 4)
data$regretaggrecoded1 = data$regretagg1+1

# Label Variables
names (data$compensationaggrecoded1) <- c("0", "100,000", "200,000", "300,000", "400,000", "500,000", "600,000", "700,000", "800,000", "900,000", "1,000,000")
names (data$regretaggrecoded1) <- c("no regret", "weak regret", "medium regret", "strong regret", "very strong regret")

#######
# Compensation
########

# what was the effect in the original study?
ci.smd(ncp=2.17, n.1=58, n.2 = 105, conf.level=0.95)

exceptioncombinedcompensationplot <- ggplot(data, aes(x=Sc3conditionl1, y=compensationaggrecoded1)) + 
  geom_violin() + geom_violin(trim=FALSE) + 
  scale_x_discrete(limits=c("Routine", "Exception")) +
  geom_jitter(shape=16, position=position_jitter(height=0.1, width=0.2)) +
  scale_color_grey() + scale_fill_grey() + 
  theme_apa() +
  theme(axis.title.x=element_blank(), 
        axis.text = element_text(size = 16),
        axis.title.y = element_text(size = 20),
        panel.grid.major = element_line(colour="grey", size = (0.1)),
        panel.grid.minor = element_line(size = (0.1), colour="grey"),
        text=element_text(family="Arial"))+
  xlab("Condition") +
  ylab("Compensation") +
  scale_y_continuous(minor_breaks = seq(-10, 15, 5)) +
  stat_summary(fun.data = mean_se, geom = "errorbar",
               position = position_dodge(width = 0.90), width = 0.2) +
  stat_summary(fun.y=mean, geom="point", shape=18,
               size=3)
exceptioncombinedcompensationplot

########
# Regret
########

exception_combined_regret_plot <- ggplot(data, aes(x = Sc3conditionl1, y = regretaggrecoded1)) +
  geom_violin(trim = FALSE) + # Duplicate geom_violin removed
  geom_jitter(shape = 16, position = position_jitter(height = 0.1, width = 0.2)) +
  scale_x_discrete(limits = c("Routine", "Exception")) +
  scale_color_grey() + 
  scale_fill_grey() + 
  theme_apa() +
  theme(
    axis.title.x = element_blank(), 
    axis.text = element_text(size = 16),
    axis.title.y = element_text(size = 20),
    panel.grid.major = element_line(colour = "grey", size = 0.1),
    panel.grid.minor = element_line(size = 0.1, colour = "grey"),
    text = element_text(family = "Arial")
  ) +
  xlab("Condition") +
  ylab("Regret") +
  stat_summary(fun.data = mean_se, geom = "errorbar",
               position = position_dodge(width = 0.9), width = 0.2) +
  stat_summary(fun.y = mean, geom = "point", shape = 18,
               size = 3)

# Print the plot
exception_combined_regret_plot

```
In the first graph, we can observe two distributions corresponding to the compensation under routine behavior conditions and exception conditions. The compensation amounts are represented on a scale from 0 to 10, which might indicate a range from $0 to $1,000,000. Under the routine behavior condition, the data distribution appears to be more concentrated, suggesting that most participants' compensation amounts were relatively close to each other, and the median seems to be near the middle value of the scale. Under the exception condition, the distribution of compensation data is broader, indicating a greater variance in the participants' assessments of the compensation amount.

The second graph illustrating the distribution of reported regret under routine and exception conditions.In the routine condition, the distribution of regret is narrower and more concentrated around the lower end of the scale, suggesting that participants generally felt less regret for the routine situation. The median regret rating, indicated by the horizontal line inside the violin plot, appears to be below the midpoint of the scale.In the exception condition, the regret distribution is wider at the top, indicating that there is a greater variance in how much regret participants felt, with more reports of higher levels of regret. The median regret rating in this condition is higher than in the routine condition, as seen by its position above the midpoint of the scale.

# Discussion

In Figure \@ref(fig:fig1), we saw that the birth rate enters a persistent decline after the Great Recession in 2007. However, one is left to wonder if this recession is really the event to demarcate this trend in birth rates, since it does not mirror the pattern of recessions prior as we have yet to see any increase in birth rates after the decline.

## Why is the birth rate declining? - Demographic insights

To investigate this idea more, we first turn to the trends among the various demographic subgroups. For the different age groups, the decreasing trend among younger age groups and increasing trend among older age groups points to the idea that women are choosing to have children later in life, an attitudinal change. Only the age 20-24 group seems to show a decrease at the 2007 mark that differs significantly from the years prior. One explanation is that it is the age when most people attend university, and more women are applying for tertiary education and pursuing education before having children. In fact, the National Center for Education Statistics (NCES) data shows a 17.6% increase in undergraduate enrollment among females from 2007 to 2011 [@nces], which coincides perfectly with the 19.1% decrease in birth rates in the 4 years after the Great Recession began in 2007.

For race and ethnicity, the birth rates for Black and White women remains fairly constant and only the Hispanic women show a significant decline in the years 2007-2011 before it plateaus to follow a similar trend to the other two races. @og offer assimilation as a possible explanation with evidence to support this notion from @tavernise and @Parrado2008.

For marital status, married and unmarried women have fairly unchanged birth rates around 2007, so it does not appear to be affected too much by the recession or other events around that time. @og suggests that the small decrease in birth rates for unmarried women is possibly due to the decline in teenage births (ages 15-19) that we saw in Figure \@ref(fig:demgraphs) plot A, since women are marrying later and therefore teenage women make up a larger proportion of the unmarried women population in their childbearing years [@marriage].

For education level, the groups do not appear to demonstrate any change in birth rates with the presence of the Great Recession, except for the no highschool degree group. Its rise and fall in birth rates may be related to the boom in the residential housing market leading up to the recession, where many subprime mortgages were given out, that is, loans to persons with low credit ratings, of which women with a lack of education i.e. no highschool degree would fall under [@insider]. Thus, more women in this group may have chosen to buy homes and start families, and then the recession hit and reversed this effect.

Thus across these demographic groups that we have examined, we note two things: 1) the the declining trend in the total female population of childbearing age is not paralleled equally across its subgroups of age, race and ethnicity, marital status and education level, but 2) specifically we note that younger age groups, Hispanic women and women with no highschool degree showed a significant decrease in birth rates of which only the no highschool degree mamas seem to be related to the Great Recession. We did mention earlier however, that there seems to be an attitudinal change among the women close to the recession - women having children later in life, getting married later, and more women attending colleges and universities - which leads into our discussion of cohort effects.

## Cohort effects - attitudes of a new generation

From Figure \@ref(fig:cohorttab), it was shown that the birth cohorts in years 1968-1982 are quite similar to each other whereas the cohorts born in years 1983-1997 seemed to pull more away with lower birth rates compared to its older counterparts. In fact, these two groupings of years define two distinct generations respectively: Generation X and Generation Y, or the "Millenials" [@gendef]. @millenial conducted a study comparing various lifestyle factors such as education, income, family and housing of Millenials to its prior generations and there are some stark differences.

Notably, women are delaying motherhood as according to a Pew Research Center analysis of population data, 2016 saw 48% of Millennial women having given birth to at least one child at ages 20 to 35 compared to the previous generation where 57% of women were already mothers in 2000 for the same age range [@milmom]. The marriage status for millenials also shows similar figures with 46% of Millennial women ages 25 to 37 being married compared to 57% of Generation X-ers for the same age range, and less millenial women are married for lower levels of education [@millenial]. Additionally, the generations are becoming more racially and ethnically diverse [@postmil] with the population of Hispanics less likely to be foreign-born [@og]. These support our arguments for the change in birth rate trends for women of the different age groups, marital status and race and ethnicity that was discussed in the previous section.

So it now begs the question, why are the attitudes between these two generations so different? We briefly touch on some historical events that surely would have influenced change in Generation Y.

## The Digital Revolution and Third Wave Feminism

The birth years of the millenial generation were characterized by several milestones in the Digital Revolution: the first mobile phone (1984), digital camera (1988) and the World Wide Web (1989) [@stfc]. When the first set of millenials would turn about 10 years old, CERN’s web browser software was released for public use (1991) and commercial dial up internet became available (1992) [@stfc]. In their teenage years to follow, smartphones, social media and high speed internet would be developed creating a tech-savvy generation like never before.

It also happens that Generation X coincides with the years of the second wave of feminism (1963 - 1980s) whereas the formative years of millenials coincide the years of the third wave of feminism [@vox]. The third wave of the feminist movement began with the Anita Hill sexual harassment case in 1991, followed by the "Year of the Woman" in 1992 where a record-breaking number of women candidates were elected to the House of Representatives and Senate (@cawp, as cited in @Dolan1998, p. 272). The third wave of feminism attempts to be distinctive from its predecessor, with some claims in popular literature of it to have a culture of 'girl power', be more open-minded, more racially diverse and inclusive than the second wave, although it is disputed whether some of these claims are exaggerated [@Snyder2008].

Therefore, the influence of technology and mass media that comes with it as well as the different ideologies behind feminism and women empowerment may have shaped the different attitudes we see between the more recent birth cohorts and older cohorts, and these changes may manifest in the decline we see in the birth rates.

## The decline of birth rates - not an effect of the Great Recession?

Back to Figure 1, we noted that the first time the birth rate dipped below minimum from previous years was in 2011, which, upon reflection after our discussion, is almost exactly 20 years after the commercial dial up internet became available and the third wave feminist movement began. 2012 is the year when the birth cohort 1988-1992 was ages 20-24, which is the first cohort to pull away significantly in birth rates to the previous ones. So are these other historical events indirectly responsible for the persistent decline in birth rates by influencing women's attitudes, and not the Great Recession? This is left as food for thought for now, more research would be needed to investigate this question.

# Conclusion

We explored the trend in birth rates in the US for the period 1980-2020 among the total child-bearing population, and demographic subgroups by age, race and ethnicity, marital status and level of education. Overall there is a decline, but the trend oftentimes remains constant for different subgroups and only declines significantly for specific groups such as younger age groups, Hispanic women and women with no highschool degree after 2007. The original paper by @og uses the Great Recession as the event to mark the beginning of the birth rate decline, but although we acknowledge some effect of the recession on decreasing birth rates, we conjecture that the lasting and continuous decline is greater explained by the shifing attitudes towards when and if to have children in the new generation, and that the decline would have been evident even if the recession had not occurred. However, if women are simply choosing to have children later and not to not have children at all, we should expect a pick back up in the birth rates in the coming decade as this group ages.

## Weaknesses and next steps

A weakness is that we only described a subset of the demographic variables that were presented in @og's paper and so the analysis could have been more complete. I do not claim to be an expert on the technlogical revolution or feminism, and we only provided these themes as potential explanatory factors for the change in attitude that is causing the decline in birth rates. More research into how these variables correlate with the different generational groups or birth cohorts is needed to make stronger statistical claims, such as by collecting data on women's beliefs and possibly constructing regression models if appropriate.

\newpage

# References

### 2024-02-12 inputs/data/osf-past-normality-regret-replication-exp1-syntax-v2-violin.R

@@ -11,9 +11,7 @@ if(!require(psych)){install.packages('psych', dependencies = TRUE)}
 if(!require(ggplot2)){install.packages('ggplot2', dependencies = TRUE)}
 if(!require(dplyr)){install.packages('dplyr', dependencies = TRUE)}
 if(!require(MBESS)){install.packages('MBESS', dependencies = TRUE)}
-if(!require(jtools)){install.packages('jtools', dependencies = TRUE)}
 
-library(jtools)
 library(MBESS)
 library(psych)
 library(ggplot2)
@@ -24,7 +22,6 @@ library(rstudioapi)
 library(jmv)
 library(Rcpp)
 
-
 # setting formatting options
 options(scipen=999.99, digits =7)
 
@@ -33,27 +30,229 @@ this.dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
 setwd(this.dir)
 
 # load our dataset 
-data <- read.csv("osf-past-normality-regret-replication-exp2-data-v2.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
+data <- read.csv("/cloud/project/inputs/data/osf-past-normality-regret-replication-exp1-data.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
 str(data)
 
-# number of participants
-NROW(na.omit(data$gender))
+# To check exclusions based on pre-registration criteria uncomment the following
+# Criteria:
+# Serious lower than 5
+# English understanding lower than 5
+# data <- data[ which(4<data$serious 
+#                     & data$engunder > 4), ]
+
 
 #Show demographics
-#Gender (1-male; 2-female)
-#set value labels
-data$gender<-factor(data$gender,levels = c(1,2,3), labels=c("Male", "Female", "Rather not say"))
+#Age
 data$gender
 table(data$gender)
-
-#Age
+#Gender
 data$age
 data$age[data$age==99] <- NA
 mean(data$age, na.rm = TRUE)
 sd(data$age, na.rm = TRUE)
 
 ###################
-# Robbery-Scenario
+# EXPERIMENT 1 (Hitchhiker-Scenario)
+
+# JAMOVI requires factors, while R imports as numeric. So, need to convert from numeric to factor.
+data$Sc1_regret<- factor(data$Sc1_regret)
[…]

### 2024-02-13 replication/scripts/fig2_replication.R
update

@@ -1,0 +1,96 @@
+#### Preamble ####
+# Purpose: Figure 2 replication
+# Author: Kimlin Chin
+# Date: 22 February 2022
+# Contact: kimlin.chin@mail.utoronto.ca
+# License: MIT
+
+## Data Sources:
+
+# Martin JA, Hamilton BE, Osterman MJK, et al. Births: Final data for 2015. 
+# National vital statistics report; vol 66, no 1. Hyattsville, MD: National Center for Health Statistics. 
+# 2017. https://www.cdc.gov/nchs/data/nvsr/nvsr66/nvsr66_01.pdf.
+# Data can be found on page 35,
+# Table 9. Birth rates, by live-birth order and race and Hispanic origin of mother: United States, 1980–2015
+# Columns: Fertility Rate, All races[years 1980-2015]
+
+# Martin JA, Hamilton BE, Osterman MJK, Driscoll AK. Births: Final Data for 2019. National
+# Vital Statistics Reports; vol 70 no 2. Hyattsville, MD: National Center for Health
+# Statistics. 2021. DOI: https://dx.doi.org/10.15620/cdc:100472.
+# Data can be found on page 13,
+# Table 1. Births and birth rates: United States, 2010–2019, and by race and Hispanic origin, United States, 2016–2019
+# Columns: Fertility Rate, All races and origins[years 2016-2018]
+
+# Hamilton BE, Martin JA, Osterman MJK. Births: Provisional data for 2020. Vital Statistics
+# Rapid Release; no 12. Hyattsville, MD: National Center for Health Statistics. May
+# 2021. DOI: https://doi.org/10.15620/cdc:104993.
+# Data can be found on page 6,
+# Table 1. Births and birth rates, by age of mother: United States, final 2019 and provisional 2020
+# Columns: 2020 Rate, 2019 Rate, Age of mother[All ages]
+
+# I crosschecked the data from these 3 reports with the data in fig_1.csv provided by 
+# the authors of the paper 
+# Kearney, Melissa S., Phillip B. Levine, and Luke Pardue. 2022. 
+# "The Puzzle of Falling US Birth Rates since the Great Recession." 
+# Journal of Economic Perspectives, 36 (1): 151-76. DOI: 10.1257/jep.36.1.151 
+# and it is the same.
+
+
[…]

### 2024-02-13 outputs/paper/paper.Rmd
update

@@
 ---
-title: "Declining Birth rates in the US: An Analysis of Potential Factors"
+title: "The Impact of Past Behavior Normality on Regret: A Replication Study"
 subtitle: "Reproduction of 'The Puzzle of Falling US Birth Rates since the Great Recession' (Kearney, Levine & Pardue, 2022)"
-author: Heng Ma
+author: Heng Ma & Bolin Shen
 thanks: "Code and data are available at: https://github.com/Brian031205/paper_2.git. A replication of various aspects in this paper are available at: ______________    "
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "Over the past 10 years, US birth rates have been dropping steadily to an all time low of 55.8 per 1,000 women in 2020.
-Kearney, Levine and Pardue (2022) analyzed this phenomenon by exploring demographic, economic, social and policy factors to little enlightenment of the cause of the decline.
-However, they did discover a correlation between the decline and the birth cohorts of mothers, and offer the conjecture of shifting priorities as the reasoning.
-We replicate the results of this study with respect to the demographic and cohort effects, and discuss alternative interpretations of the data considering the impact of the feminist movement and the evolution of technology and media.
-The results were consistent with the initial study."
+abstract: "Norm theory (Kahneman & Miller 1986) described a tendency for people to associate stronger regret with a negative outcome when it is a result of an exception (abnormal behavior) compared to when it is a result of routine (normal behavior).
+In this paper, we conducted a replication of one classic experiments on past behavior exception/routine contrasts.
+We successfully replicated Kahneman and Miller’s (1986) experiments with the classic hitchhiker-scenario (Part 1) and car accident-scenario (Part 2).
+In both cases, participants examined negative outcomes and tended to indicate a protagonist who deviated from own past behavior as more regretful than another who followed routine.
+In robbery scenario experiments using compensatory measures (Part 3), there is no significant difference between special and regular situations in the compensation comparison of robbery victims.
+However, it is showed that robberies in exceptional circumstances are considered more regrettable than robberies in regular circumstances.
+We discuss implications for current and future research."
 output:
@@
 
-<!-- Background / Broader Context -->
-Birth rate is the number of live births per thousand in the population by year [@dict].
-Birth rate is important since it is an indicator of population growth and determines the age structure of populations, which has various implications for the economy and society [@CLELAND2008364].
-When the birth rate is too high or low, it affects all subgroups of the population in different ways, for example, a large number of new births will be a burden on the adult population to support them, and when this group ages to be elderly they will be a burden on governments to support them [@CLELAND2008364].
+Regret is a pervasive and intense negative emotion, plays a pivotal role in shaping individuals' daily experiences.
+The recognition that other decisions or circumstances could have led to more favorable outcomes triggers regret (Byrne, 2016; Roese & Epstude, 2017; Kahneman & Tversky, 1982).
+Influential work by Kahneman and Miller (1986) in norm theory has highlighted the connection between abnormal behavior, heightened counterfactual thinking, and intensified feelings of regret.
+Their discussion of the concept of normality included experimental research illustrating various factors influencing the perception and interpretation of normality (Feldman & Albarracín, 2017; Feldman, 2018).
+This approach often employed past behavior as a reference point to assess the normality of an individual's actions.
 
-There are many possible factors that can affect birth rate.
-Recessions have been proposed to result in a drop in birth rates, specifically the Great Recession of 2007 [@pew; @focus; @econ].
-@becker proposes an economic framework for fertility, where the "demand for children" is driven by the satisfaction that children brings to people weighted against factors such as preferences (e.g.
-religion, race, age), income, time, expenditure and means to support each child and other considerations (e.g.
-contraception, marriage).
-@og base their analysis of the declining US birth rates on these factors.
+Illustrating the effect of past behavior normality, Kahneman and Miller (1986) presented a thought experiment involving a car accident.
+Mr.
+White, accustomed to driving home along the same route daily, deviated from his routine one day and found himself in an accident.
+82% of participants believed that Mr.
+White would experience more regret than those involved in similar accidents on his usual route.
+This bias referenced as the "exceptionality effect," serves as a basis for this study, emphasizing the regularity of past behavior.
 
-These analyses however, did not provide sufficient evidence to explain the decline in birth rates, with the authors stating "For any factor to have explained much of that decline, it would have had to change dramatically around the same time." [@og].
-@og then went on to find a correlation between the recent birth cohorts of mothers and the substantial decline.
-In our reproduction of their paper, we build on their 'shifting priorities' explanation of the cohort effects by considering the other notable events that potentially may have influenced the nature of parenting of these cohorts, namely the Third wave of Feminism and the second half of the Digital Revolution.
+Kahneman and Miller's (1986) norm theory provides a psychological classification of events as normal or abnormal, assuming that abnormal events elicit stronger reactions because of the increased accessibility of normal events in memory.
+This high availability of psychologically simulated alternatives intensifies regret over negative outcomes.
+Everyday behaviors, more easily recalled and psychologically emphasized, enable individuals to vividly imagine counterfactual scenarios in exceptional circumstances, thus amplifying the regret associated with deviations from past behavior (Kahneman & Miller 1986).
 
-We replicate the paper by @og with a focus on the following research questions:
+This study aims to directly replicate three experiments that demonstrated the exceptionality effect, representing early investigations into the influence of past behavior normality on regret.
+Specifically, we investigate hitchhiking and car accident scenarios from Kahneman and Miller (1986) in the first two part and replicate Miller and McFarland's (1986) Experiment 1, featuring a robbery scenario, in the part three to test norm theory.
+Although there is no documented direct replication of these experiments, the existing citation impact of Kahneman and Miller's work emphasizes the significance of a reexamination and detailed explanation of these effects.
+The study's findings are expected to contribute to a better understanding of the general nature of the exceptionality effect and the intricate influence of past behavior normality on regret.
 
-* What is the trend in US birth rates over the period 1980-2020?
-* How does the trend in birth rates vary with demographic factors such as age, race, education and marital status?
-* How do the birth rates vary for different cohorts of mothers by their birth year?
-
-While the original paper used Stata [@citeStata] for data processing and analysis in its replication package, we use R [@citeR] for all data wrangling and analysis and R packages tidyverse [@tidy], ggthemes [@ggthemes], ggprism [@ggprism] and patchwork [@patchwork] to produce the figures, kableExtra [@kableExtra] to produce the tables and haven [@haven] to read the dta files.
+we use R [@citeR] for all data wrangling and analysis and R packages tidyverse [@tidy], ggthemes [@ggthemes], ggprism [@ggprism] and patchwork [@patchwork] to produce the figures, kableExtra [@kableExtra] to produce the tables.
 
 # Data
+The data we used in this paper are obtained from the paper titled “The impact of past behavior normality on regret: replication and extension of three experiments of the exceptionality effect” (Kutscher & Feldman 2022).
+We first cleaned the data, and then analyzed them by the open-source statistical programming language R (R Core Team 2022).
+We used functionalities from tidyverse (Wickham et al.
+2019), ggplot2 (Wickham 2016), dplyr (Wickham et al.
+2022), readr (Wickham, Hester, and Bryan 2022), tibble (Müller and Wickham 2022), janitor (Firke 2021) and knitr (Xie 2014).
 
-## Data Source and Methodology
+## Proportion for perceived regret, injunctive social norms, descriptive social norms, and negative affect
+The first figure presents an overview of the perceived regret, injunctive social norms, descriptive social norms, and negative affect proportions associated with the behaviors of Mr.
+Smith (routine) and Mr.
+Jones (exception).
+In terms of regret, the figure highlights an obvious contrast, with Mr.
+Smith at 7.9% and Mr.
+Jones at a much higher 92.1%.
+A similar pattern emerges for injunctive social norms, where Mr.
+Smith scores 4.7%, contrasting sharply with Mr.
+Jones at 95.3%.
+Descriptive social norms show an obvious preference for routine behavior, as reflected in Mr.
+Smith's 90.6%, while Mr.
+Jones lags behind at 9.4%.
+Negative affect follows a parallel trend, with Mr.
+Smith at 7.3% and Mr.
+Jones at 92.7%.
+This figure vividly illustrates the remarkable differences in perceptions between routine and exceptional behaviors, emphasizing the notable impact on regret and social norm evaluations.
 
-The data on birth rates per 1,000 women ages 15-44 across all races and demographic population subgroups are obtained from the National Vital Statistics Reports for the years 2015, 2019 and 2020 [@2015birth; @2019birth; @2020birth], which they collected from birth certificates registered in all  US states and the District of Colombia.
-Aggregated data on the number of births for 6 different 5-year cohorts of mothers by their age and birth year is provided by @og, calculated using public birth microdata across the period 1980-1989 from the NBER Natality Database and NCHS microdata from the period 1990-2019 [@nchsR; @nber].
-The birth cohorts span the years 1968-1997.
-The NCHS microdata is restricted use and requires an application to be obtained.
-We simply used the data that was supplied in the replication packageby @og.
+## Proportion for perceived regret and luck
+Figure 2, titled "Proportions for Perceived Regret and Luck," provides a clear visual representation of the contrasts in perceived regret and luck between the behaviors of Mr.
+Adams (routine) and Mr.
+White (exception).
+The data shows a large difference in regret percentages, with Mr.
+Adams scoring at 19%, while Mr.
+White gets a significantly higher 81%.
+The figure also reveals the aspect of luck, where Mr.
+Adams has a proportion of 33.3%, distinctly lower than Mr.
+White, who gets a larger share at 66.7%.
+This figure effectively demonstrates the difference in perceptions regarding regret and luck associated with routine and exceptional behaviors, revealing the dynamic relationship between these two factors in the participants' evaluations of the scenarios described.
 
-Single-age population counts, among all races from 1969-2019 and by race and Hispanic origin from 1990-2019, is obtained from the CDC SEER database [@nci].
-
-## Attributes
-
-The birth data consists of variables for the year, and birth rates for each population subgroup.
-For example, for race there are 3 variables: birth rate for whites, birth rate for blacks and birth rate for hispanics.
-There is birth rate data for 4 population subgroups - age (5-year groupings), race and ethnicity, marital status and mother's level of education.
-We show a glimpse of the data with variables for race and marital status in table \@ref(tab:birthdata).
+## Victim compensation and regret (with exception conditions combined)
+The last figure titled "Victim Compensation and Regret (with Exception Conditions Combined)," presents an overview of two experiments, each with routine and exception conditions.
+The top two plots correspond to Experiment 1, while the bottom two plots represent Experiment 2, with error bars indicating standard errors.
+The scales range from 1 to 5 for compensation and 1 to 10 for regret.
+In Experiment 1, the routine condition includes 114 participants, with standard deviations of 2.37 for compensation and 1.12 for regret.
+The self-produced exception condition involves 115 participants, with standard deviations of 2.46 for compensation and 0.80 for regret.
+For the other-produced exception condition (113 participants), standard deviations are 2.76 for compensation and 0.94 for regret.
+Combining all exception conditions (228 participants) gets standard deviations of 2.61 for compensation and 0.87 for regret.
+In Experiment 2, routine conditions involve 116 participants, with standard deviations of 3.40 for compensation and 1.19 for regret.
+The self-produced exception condition (110 participants) shows standard deviations of 3.50 for compensation and 0.67 for regret.
+The other-produced exception condition (116 participants) has standard deviations of 3.66 for compensation and 0.98 for regret.
+Combining all exception conditions (226 participants) results in standard deviations of 3.57 for compensation and 0.86 for regret.
+These detailed data provide an understanding of victim compensation and regret across various experimental conditions.
 

### 2024-02-13 outputs/paper/paper.Rmd
1: Update readme file withe a simple introduction for the project. 2:knit to the pdf file 3:add the figures in replication folder 4:fix some mistakes for the codes

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-knitr::opts_chunk$set(fig.pos = "h", out.extra = "")
 
@@
 if(!require(MBESS)){install.packages('MBESS', dependencies = TRUE)}
+if(!require(knitr)){install.packages('knitr', dependencies = TRUE)}
+if(!require(jtools)){install.packages('jtools', dependencies = TRUE)}
 
+library(jtools)
+library(knitr)
 library(MBESS)
@@
 
-```{r birthdata, echo=FALSE}
-# Makes table of Ten rows of the birth rate data for years 1980-2020
-birth_data %>%
-  select(year, brate_all, brate_whitenh, brate_blacknh, brate_hisp, brate_unmarried, brate_married) %>%
-  filter(year > 1989) %>%
-  slice(1:10) %>%
-  kable(
-    caption = "Ten rows of the birth rate data for years 1980-2020",
-    col.names = c("Year", "All races", "White, non-Hispanic", "Black, non-Hispanic", "Hispanic", "Unmarried", "Married"),
-    digits = 1,
-    booktabs = TRUE, 
-    linesep = ""
-  )
-```
-
 For the aggregated cohort birth data, the variables are mother's age, birth cohort, number of births and cumulative births.
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-# Process population data to aggregate by cohort
-fig5_pop_data1 <- pop_cohort_data %>%
-  select(1:32) %>%
-  select(-c(stname)) %>%
-  group_by(year) %>%
-  summarise(across(everything(),sum)) # https://stackoverflow.com/questions/1660124/how-to-sum-a-variable-by-group
-
-fig5_pop_data1 <- fig5_pop_data1 %>%
-  pivot_longer(-c(year), names_to = "mage", values_to = "pop")
-
-fig5_pop_data1 <- fig5_pop_data1 %>%
-  mutate(mage = as.numeric(str_replace(mage, "fem", ""))) %>%
-  mutate(cohort = year - mage) %>%
-  mutate(cohort2 = case_when(cohort >= 1968 & cohort <=1972 ~ 1,
-                             cohort >= 1973 & cohort <=1977 ~ 2,
-                             cohort >= 1978 & cohort <=1982 ~ 3,
-                             cohort >= 1983 & cohort <=1987 ~ 4,
-                             cohort >= 1988 & cohort <=1992 ~ 5,
-                             cohort >= 1993 & cohort <=1997 ~ 6))
-
-fig5_pop_data1 <- na.omit(fig5_pop_data1)
-fig5_pop_data1 <- fig5_pop_data1 %>%
-  group_by(cohort2, mage) %>%
-  summarise(pop = sum(pop))
-
-# Join dataframes for population and birth data
-birth_cohort_data <- full_join(birth_cohort_data, fig5_pop_data1, by=c("mage", "cohort2")) %>%
-  mutate(cohort_years = case_when(cohort2 == 1 ~ "1968-1972",
-                                  cohort2 == 2 ~ "1973-1977",
-                                  cohort2 == 3 ~ "1978-1982",
-                                  cohort2 == 4 ~ "1983-1987",
-                                  cohort2 == 5 ~ "1988-1992",
-                                  cohort2 == 6 ~ "1993-1997"))
-
-```
-
-```{r birthcohort, echo=FALSE}
-# Make table of First ten rows of the birth cohort data for 6 birth cohorts
-birth_cohort_data %>%
-  select(2,6,1,3,4,5) %>%
-  slice(1:10) %>%
-  kable(
-    caption = "First ten rows of the birth cohort data for 6 birth cohorts",
-    col.names = c("Cohort #", "Birth Years", "Mother's age", "Number of births", "Cumulative births", "Population Count"),
-    digits = 1,
-    booktabs = TRUE, 
-    linesep = ""
-  )
-```
-
 # Results
@@
 ## Overall evaluation in regre depending hitchhiker scenario
-
-```{r table1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Rates for Regret, Social norms and Negative affect"}
-
-Orginal_table <- read.csv("/cloud/project/inputs/data/Ex1.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
-print(Orginal_table)
-```
 
@@
 ```{r fig1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Rates for Regret, Social norms and Negative affect"}
-data1 <- read.csv("/cloud/project/inputs/data/Ex1.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
-str(data)
+data1 <- read.csv("/cloud/project/inputs/data/Ex1.csv", header = TRUE, stringsAsFactors = FALSE)
 
@@
 
-ggplot(long_data, aes(x = Category, y = Percentage, fill = Group)) +
+fig1 <- ggplot(long_data, aes(x = Category, y = Percentage, fill = Group)) +
   geom_bar(stat = 'identity', position = 'dodge') +
@@
   )
-print(ggplot)
+fig1
 ```
 
-## Demographic Factors
+## Car accident scenario
 
-Figure \@ref(fig:fig2) is a replication of Figure 2 in the original paper, and shows the Proportions for perceived regret and luck.This graph compares the relationship between regret and luck, with 81% of 277 respondents believing that Mr.
-White, who had just changed to a new route home and experienced an accident, would feel more regretful.
-```{r table1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Rates for Regret, Social norms and Negative affect"}
-
-table2 <- read.csv("/cloud/project/inputs/data/table2.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
-print(table2)
-```
- As for luck, Mr.
+Figure 2 is a replication of Figure 2 in the original paper, and shows the Proportions for perceived regret and luck.This graph compares the relationship between regret and luck, with 81% of 277 respondents believing that Mr.
+White, who had just changed to a new route home and experienced an accident, would feel more regretful.As for luck, Mr.
 White (exception) was considered less lucky than Mr.
@@
 
-```{r fig2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Trends in Birth Rates by Population Subgroup"}
+```{r fig2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Proportions for perceived regret and luck"}
 
@@
 
-# Clean up the data by removing unnecessary 'Counts' columns and renaming the remaining
 df_clean <- df[, -c(2, 4)]
@@
 
-# Melt the data frame to have 'Category', 'Variable', and 'Value' columns
 df_melted <- melt(df_clean, id.vars = "Category", variable.name = "Outcome", value.name = "Percentage")
 
-# Plot with ggplot2
-p <- ggplot(df_melted, aes(x = Outcome, y = Percentage, fill = Category)) +
+fig2 <- ggplot(df_melted, aes(x = Outcome, y = Percentage, fill = Category)) +
   geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
@@
 
-# Print the plot
-print(p)
+fig2
 ```
 
-```{r proportions, echo=FALSE, warning=FALSE, message=FALSE}
-this.dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
-setwd(this.dir)
-
-# load our dataset 
-data <- read.csv("/cloud/project/inputs/data/osf-past-normality-regret-replication-exp1-data.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
-str(data)
-
-data$Sc1_regret<- factor(data$Sc1_regret)
-data$sc1_socnorms1<- factor(data$sc1_socnorms1)
-data$sc1_socnorms2<- factor(data$sc1_socnorms2)
-data$sc1_combinednorms<- factor(data$sc1_combinednorms)
-
-# Let's label the values better, so it's easier to understand the output.
-data$Sc1_regret <- ordered(data$Sc1_regret, levels = c(1,2), labels = c("Exception Jones", "Routine Smith"))
-data$sc1_socnorms1 <- ordered(data$sc1_socnorms1, levels = c(1,2), labels = c("Exception Jones", "Routine Smith"))
-data$sc1_socnorms2 <- ordered(data$sc1_socnorms2, levels = c(1,2), labels = c("Exception Jones", "Routine Smith"))
-data$sc1_combinednorms <- ordered(data$sc1_combinednorms, levels = c(1,2), labels = c("Exception Jones", "Routine Smith"))
-
-data$Sc3condition1 <- 0
-data$compensationagg1 <- 0
-data$regretagg1 <- 0
-for (i in 1:nrow(data)){
-  if (!is.na(data$Sc3_C1_text[i])){
-    data$Sc3condition1[i] <- 1
-    data$compensationagg1[i] <- data$sc3_c1_compensation[i]
-    data$regretagg1[i] <- data$sc3_c1_regret[i]
-  }
-  else if (!is.na(data$Sc3_C2_text[i])){
-    data$Sc3condition1[i] <- 2
-    data$compensationagg1[i] <- data$sc3_c2_compensation[i]
-    data$regretagg1[i] <- data$sc3_c2_regret[i]
-  }
-  else if (!is.na(data$Sc3_C3_text[i])){
-    data$Sc3condition1[i] <- 2
-    data$compensationagg1[i] <- data$sc3_c3_compensation[i]
-    data$regretagg1[i] <- data$sc3_c3_regret[i]
-  }
-  else {
-    data$Sc3condition1[i] <- NA
-    data$compensationagg1[i] <- NA
-    data$regretagg1[i] <- NA
-  }
-}
-
-#value labels
-data$Sc3conditionl1<-factor(data$Sc3condition1,levels = c(1,2), labels=c("Routine", "Exception"))
-
-# let's have a look at this
-table(data$Sc3conditionl1)
-
-#Adjust Copensation-Scale to original paper (Values from 0 to 10, instead of 1 to 11)
-data$compensationaggrecoded1 = data$compensationagg1-1
-
-#Adjust Regret Scale (Values 1 to 5, indead of 0 to 4)
-data$regretaggrecoded1 = data$regretagg1+1
-
-# Label Variables
-names (data$compensationaggrecoded1) <- c("0", "100,000", "200,000", "300,000", "400,000", "500,000", "600,000", "700,000", "800,000", "900,000", "1,000,000")
-names (data$regretaggrecoded1) <- c("no regret", "weak regret", "medium regret", "strong regret", "very strong regret")
-
-#######
-# Compensation
-########
-
-# what was the effect in the original study?
-ci.smd(ncp=2.17, n.1=58, n.2 = 105, conf.level=0.95)
-
-exceptioncombinedcompensationplot <- ggplot(data, aes(x=Sc3conditionl1, y=compensationaggrecoded1)) + 
-  geom_violin() + geom_violin(trim=FALSE) + 
-  scale_x_discrete(limits=c("Routine", "Exception")) +
-  geom_jitter(shape=16, position=position_jitter(height=0.1, width=0.2)) +
-  scale_color_grey() + scale_fill_grey() + 
-  theme_apa() +
-  theme(axis.title.x=element_blank(), 
-        axis.text = element_text(size = 16),
-        axis.title.y = element_text(size = 20),
-        panel.grid.major = element_line(colour="grey", size = (0.1)),
-        panel.grid.minor = element_line(size = (0.1), colour="grey"),
-        text=element_text(family="Arial"))+
-  xlab("Condition") +
-  ylab("Compensation") +
-  scale_y_continuous(minor_breaks = seq(-10, 15, 5)) +
-  stat_summary(fun.data = mean_se, geom = "errorbar",
-               position = position_dodge(width = 0.90), width = 0.2) +
-  stat_summary(fun.y=mean, geom="point", shape=18,
-               size=3)
-exceptioncombinedcompensationplot
-
-########
-# Regret
-########
-
-exception_combined_regret_plot <- ggplot(data, aes(x = Sc3conditionl1, y = regretaggrecoded1)) +
-  geom_violin(trim = FALSE) + # Duplicate geom_violin removed
-  geom_jitter(shape = 16, position = position_jitter(height = 0.1, width = 0.2)) +
-  scale_x_discrete(limits = c("Routine", "Exception")) +
-  scale_color_grey() + 
-  scale_fill_grey() + 
-  theme_apa() +
-  theme(
-    axis.title.x = element_blank(), 
-    axis.text = element_text(size = 16),
-    axis.title.y = element_text(size = 20),
-    panel.grid.major = element_line(colour = "grey", size = 0.1),
-    panel.grid.minor = element_line(size = 0.1, colour = "grey"),
-    text = element_text(family = "Arial")
-  ) +
-  xlab("Condition") +
-  ylab("Regret") +
-  stat_summary(fun.data = mean_se, geom = "errorbar",
-               position = position_dodge(width = 0.9), width = 0.2) +
-  stat_summary(fun.y = mean, geom = "point", shape = 18,
-               size = 3)
-
-# Print the plot
-exception_combined_regret_plot
-
-```
-In the first graph, we can observe two distributions corresponding to the compensation under routine behavior conditions and exception conditions.
+Figure 3 In the first graph, we can observe two distributions corresponding to the compensation under routine behavior conditions and exception conditions.
 The compensation amounts are represented on a scale from 0 to 10, which might indicate a range from $0 to $1,000,000.
@@
 
+```{r fig3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="compensation and reported regret "}
+
+data <- read.csv("/cloud/project/inputs/data/osf-past-normality-regret-replication-exp1-data.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
+data$Sc1_regret <- factor(data$Sc1_regret, levels = c(1,2), ordered = TRUE, labels = c("Exception Jones", "Routine Smith"))
+data$sc1_socnorms1 <- factor(data$sc1_socnorms1, levels = c(1,2), ordered = TRUE, labels = c("Exception Jones", "Routine Smith"))
+data$sc1_socnorms2 <- factor(data$sc1_socnorms2, levels = c(1,2), ordered = TRUE, labels = c("Exception Jones", "Routine Smith"))
+data$sc1_combinednorms <- factor(data$sc1_combinednorms, levels = c(1,2), ordered = TRUE, labels = c("Exception Jones", "Routine Smith"))
+
+data$Sc3condition1 <- NA
+data$compensationagg1 <- NA
+data$regretagg1 <- NA
+
+data <- transform(data, Sc3condition1 = ifelse(!is.na(Sc3_C1_text), 1, ifelse(!is.na(Sc3_C2_text) | !is.na(Sc3_C3_text), 2, NA)),
+                  compensationagg1 = ifelse(!is.na(Sc3_C1_text), sc3_c1_compensation, ifelse(!is.na(Sc3_C2_text), sc3_c2_compensation, ifelse(!is.na(Sc3_C3_text), sc3_c3_compensation, NA))),
+                  regretagg1 = ifelse(!is.na(Sc3_C1_text), sc3_c1_regret, ifelse(!is.na(Sc3_C2_text), sc3_c2_regret, ifelse(!is.na(Sc3_C3_text), sc3_c3_regret, NA))))
+
+data$Sc3conditionl1 <- factor(data$Sc3condition1, levels = c(1, 2), labels = c("Routine", "Exception"))
+
+data$compensationaggrecoded1 <- data$compensationagg1 - 1
+data$regretaggrecoded1 <- data$regretagg1 + 1
+
+ggplot(data, aes(x = Sc3conditionl1, y = compensationaggrecoded1)) + 
+  geom_violin(trim = FALSE) + 
+  geom_jitter(shape = 16, position = position_jitter(height = 0.1, width = 0.2)) + 
+  theme_apa() +
+  labs(x = "Condition", y = "Compensation") +
+  scale_y_continuous(minor_breaks = seq(-10, 15, 5)) +
+  stat_summary(fun.data = mean_se, geom = "errorbar", position = position_dodge(width = 0.9), width = 0.2) +
+  stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3)
+
+ggplot(data, aes(x = Sc3conditionl1, y = regretaggrecoded1)) + 
+  geom_violin(trim = FALSE) + 
+  geom_jitter(shape = 16, position = position_jitter(height = 0.1, width = 0.2)) + 
+  theme_apa() +
+  labs(x = "Condition", y = "Regret") +
+  stat_summary(fun.data = mean_se, geom = "errorbar", position = position_dodge(width = 0.9), width = 0.2) +
+  stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3)
+
+```
+
 # Discussion
-
-In Figure \@ref(fig:fig1), we saw that the birth rate enters a persistent decline after the Great Recession in 2007.
-However, one is left to wonder if this recession is really the event to demarcate this trend in birth rates, since it does not mirror the pattern of recessions prior as we have yet to see any increase in birth rates after the decline.
 
@@
 
-To investigate this idea more, we first turn to the trends among the various demographic subgroups.
-For the different age groups, the decreasing trend among younger age groups and increasing trend among older age groups points to the idea that women are choosing to have children later in life, an attitudinal change.
-Only the age 20-24 group seems to show a decrease at the 2007 mark that differs significantly from the years prior.
-One explanation is that it is the age when most people attend university, and more women are applying for tertiary education and pursuing education before having children.
-In fact, the National Center for Education Statistics (NCES) data shows a 17.6% increase in undergraduate enrollment among females from 2007 to 2011 [@nces], which coincides perfectly with the 19.1% decrease in birth rates in the 4 years after the Great Recession began in 2007.
-
-For race and ethnicity, the birth rates for Black and White women remains fairly constant and only the Hispanic women show a significant decline in the years 2007-2011 before it plateaus to follow a similar trend to the other two races.
-@og offer assimilation as a possible explanation with evidence to support this notion from @tavernise and @Parrado2008.
-
-For marital status, married and unmarried women have fairly unchanged birth rates around 2007, so it does not appear to be affected too much by the recession or other events around that time.
-@og suggests that the small decrease in birth rates for unmarried women is possibly due to the decline in teenage births (ages 15-19) that we saw in Figure \@ref(fig:demgraphs) plot A, since women are marrying later and therefore teenage women make up a larger proportion of the unmarried women population in their childbearing years [@marriage].
-
-For education level, the groups do not appear to demonstrate any change in birth rates with the presence of the Great Recession, except for the no highschool degree group.
-Its rise and fall in birth rates may be related to the boom in the residential housing market leading up to the recession, where many subprime mortgages were given out, that is, loans to persons with low credit ratings, of which women with a lack of education i.e.
-no highschool degree would fall under [@insider].
-Thus, more women in this group may have chosen to buy homes and start families, and then the recession hit and reversed this effect.
-
-Thus across these demographic groups that we have examined, we note two things: 1) the the declining trend in the total female population of childbearing age is not paralleled equally across its subgroups of age, race and ethnicity, marital status and education level, but 2) specifically we note that younger age groups, Hispanic women and women with no highschool degree showed a significant decrease in birth rates of which only the no highschool degree mamas seem to be related to the Great Recession.
-We did mention earlier however, that there seems to be an attitudinal change among the women close to the recession - women having children later in life, getting married later, and more women attending colleges and universities - which leads into our discussion of cohort effects.
-
 ## Cohort effects - attitudes of a new generation
-
-From Figure \@ref(fig:cohorttab), it was shown that the birth cohorts in years 1968-1982 are quite similar to each other whereas the cohorts born in years 1983-1997 seemed to pull more away with lower birth rates compared to its older counterparts.
-In fact, these two groupings of years define two distinct generations respectively: Generation X and Generation Y, or the "Millenials" [@gendef].
-@millenial conducted a study comparing various lifestyle factors such as education, income, family and housing of Millenials to its prior generations and there are some stark differences.
-
-Notably, women are delaying motherhood as according to a Pew Research Center analysis of population data, 2016 saw 48% of Millennial women having given birth to at least one child at ages 20 to 35 compared to the previous generation where 57% of women were already mothers in 2000 for the same age range [@milmom].
-The marriage status for millenials also shows similar figures with 46% of Millennial women ages 25 to 37 being married compared to 57% of Generation X-ers for the same age range, and less millenial women are married for lower levels of education [@millenial].
-Additionally, the generations are becoming more racially and ethnically diverse [@postmil] with the population of Hispanics less likely to be foreign-born [@og].
-These support our arguments for the change in birth rate trends for women of the different age groups, marital status and race and ethnicity that was discussed in the previous section.
-
-So it now begs the question, why are the attitudes between these two generations so different? We briefly touch on some historical events that surely would have influenced change in Generation Y.
 
@@
 
-The birth years of the millenial generation were characterized by several milestones in the Digital Revolution: the first mobile phone (1984), digital camera (1988) and the World Wide Web (1989) [@stfc].
-When the first set of millenials would turn about 10 years old, CERN’s web browser software was released for public use (1991) and commercial dial up internet became available (1992) [@stfc].
-In their teenage years to follow, smartphones, social media and high speed internet would be developed creating a tech-savvy generation like never before.
-
-It also happens that Generation X coincides with the years of the second wave of feminism (1963 - 1980s) whereas the formative years of millenials coincide the years of the third wave of feminism [@vox].
-The third wave of the feminist movement began with the Anita Hill sexual harassment case in 1991, followed by the "Year of the Woman" in 1992 where a record-breaking number of women candidates were elected to the House of Representatives and Senate (@cawp, as cited in @Dolan1998, p.
-272).
-The third wave of feminism attempts to be distinctive from its predecessor, with some claims in popular literature of it to have a culture of 'girl power', be more open-minded, more racially diverse and inclusive than the second wave, although it is disputed whether some of these claims are exaggerated [@Snyder2008].
-
-Therefore, the influence of technology and mass media that comes with it as well as the different ideologies behind feminism and women empowerment may have shaped the different attitudes we see between the more recent birth cohorts and older cohorts, and these changes may manifest in the decline we see in the birth rates.
-
 ## The decline of birth rates - not an effect of the Great Recession?
-
-Back to Figure 1, we noted that the first time the birth rate dipped below minimum from previous years was in 2011, which, upon reflection after our discussion, is almost exactly 20 years after the commercial dial up internet became available and the third wave feminist movement began.
-2012 is the year when the birth cohort 1988-1992 was ages 20-24, which is the first cohort to pull away significantly in birth rates to the previous ones.
-So are these other historical events indirectly responsible for the persistent decline in birth rates by influencing women's attitudes, and not the Great Recession?
-This is left as food for thought for now, more research would be needed to investigate this question.
 
@@
 
-We explored the trend in birth rates in the US for the period 1980-2020 among the total child-bearing population, and demographic subgroups by age, race and ethnicity, marital status and level of education.
-Overall there is a decline, but the trend oftentimes remains constant for different subgroups and only declines significantly for specific groups such as younger age groups, Hispanic women and women with no highschool degree after 2007.
-The original paper by @og uses the Great Recession as the event to mark the beginning of the birth rate decline, but although we acknowledge some effect of the recession on decreasing birth rates, we conjecture that the lasting and continuous decline is greater explained by the shifing attitudes towards when and if to have children in the new generation, and that the decline would have been evident even if the recession had not occurred.
-However, if women are simply choosing to have children later and not to not have children at all, we should expect a pick back up in the birth rates in the coming decade as this group ages.
-
 ## Weaknesses and next steps
-
-A weakness is that we only described a subset of the demographic variables that were presented in @og's paper and so the analysis could have been more complete.
-I do not claim to be an expert on the technlogical revolution or feminism, and we only provided these themes as potential explanatory factors for the change in attitude that is causing the decline in birth rates.
-More research into how these variables correlate with the different generational groups or birth cohorts is needed to make stronger statistical claims, such as by collecting data on women's beliefs and possibly constructing regression models if appropriate.
 

### 2024-02-14 outputs/paper/paper.qmd
update

--- 2024-02-13 outputs/paper/paper.Rmd
@@
 
-Regret is a pervasive and intense negative emotion, plays a pivotal role in shaping individuals' daily experiences.
+Regret, a pervasive and intense negative emotion, plays a pivotal role in shaping individuals' daily experiences.
 The recognition that other decisions or circumstances could have led to more favorable outcomes triggers regret (Byrne, 2016; Roese & Epstude, 2017; Kahneman & Tversky, 1982).
@@
 This study aims to directly replicate three experiments that demonstrated the exceptionality effect, representing early investigations into the influence of past behavior normality on regret.
-Specifically, we investigate hitchhiking and car accident scenarios from Kahneman and Miller (1986) in the first two part and replicate Miller and McFarland's (1986) Experiment 1, featuring a robbery scenario, in the part three to test norm theory.
+Specifically, we investigate hitchhiking and car accident scenarios from Kahneman and Miller (1986) in the first two part and replicate Miller and McFarland's (1986) Experiment 1, featuring a robbery scenario, in the third part to test norm theory.
+There are three research questions: for hitchhiking scenarios from Kahneman and Miller (1986), we investigated the proportion for perceived regret, injunctive social norms, descriptive social norms, and negative affect.
+For car accident scenario, we studied the proportion for perceived regret and luck.
+And for the scenario of robbery, we investigated victim compensation and regret (with exception conditions combined).
 Although there is no documented direct replication of these experiments, the existing citation impact of Kahneman and Miller's work emphasizes the significance of a reexamination and detailed explanation of these effects.

### 2024-02-15 outputs/paper/Paper_2.qmd
1:Organize all the files with the new method of analize of data and new topic. 2:delect the orginal paper, create paper_2 3:add the paragraph of Discussion parts 4:remove usless references.

---
title: "The Impact of Past Behavior Normality on Regret: A Replication Study"
subtitle: "Reproduction of 'The relationship between the three elements: new experiences, luck, and compensation."
author: Heng Ma & Bolin Shen
thanks: "Code and data are available at: https://github.com/Brian031205/paper_2.git. A replication of various aspects in this paper are available at:https://www.socialsciencereproduction.org/reproductions/bee4097e-9c14-4c32-baef-cf9f3fa61a2b/index"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Norm theory (Kahneman & Miller 1986) described a tendency for people to associate stronger regret with a negative outcome when it is a result of an exception (abnormal behavior) compared to when it is a result of routine (normal behavior). In this paper, we conducted a replication of one classic experiments on past behavior exception/routine contrasts. We successfully replicated Kahneman and Miller’s (1986) experiments with the classic hitchhiker-scenario (Part 1) and car accident-scenario (Part 2). In both cases, participants examined negative outcomes and tended to indicate a protagonist who deviated from own past behavior as more regretful than another who followed routine. In robbery scenario experiments using compensatory measures (Part 3),However, these analyses do not describe the relationship between the three elements: new experiences, luck, and compensation. This article, therefore, focuses on emphasizing which of these elements more significantly affects how individuals measure the extent of their regret,and then analyze the role of various factors from a psychological perspective. "
format: pdf
toc: TRUE
bibliography: references.bib
---

```{r setup, include=FALSE}

# Load libraries
if(!require(Rcpp)){install.packages('Rcpp', dependencies = TRUE)}
if(!require(jmv)){install.packages('jmv', repos=c('https://repo.jamovi.org', 'https://cran.r-project.org'))}
if(!require(jmv)){install.packages('jmv', dependencies = TRUE)}
if(!require(rstudioapi)){install.packages('rstudioapi', dependencies = TRUE)}
if(!require(effsize)){install.packages('effsize', dependencies = TRUE)}
if(!require(Hmisc)){install.packages('effsize', dependencies = TRUE)}
if(!require(psych)){install.packages('psych', dependencies = TRUE)}
if(!require(ggplot2)){install.packages('ggplot2', dependencies = TRUE)}
if(!require(dplyr)){install.packages('dplyr', dependencies = TRUE)}
if(!require(MBESS)){install.packages('MBESS', dependencies = TRUE)}
if(!require(knitr)){install.packages('knitr', dependencies = TRUE)}
if(!require(jtools)){install.packages('jtools', dependencies = TRUE)}

library(jtools)
library(knitr)
library(MBESS)
library(psych)
library(ggplot2)
library(dplyr) 
library(Hmisc)
library(effsize)
library(rstudioapi)
library(jmv)
library(Rcpp)

```

# Introduction

Regret, a profound and often debilitating emotion, plays a significant role in the fabric of human experience, weaving its way through our decisions, actions, and reflections. This emotional response, characterized by a sense of loss or disappointment over what might have been, emerges from the recognition that different decisions or circumstances could have led to more favorable outcomes. While regret is a universal experience, its intensity and impact can vary widely among individuals and situations. In this article, we delve into the complex nature of regret, exploring its triggers, consequences, and the psychological underpinnings that shape this emotional landscape. Through a focused examination of three distinct scenarios—Hitchhiker, Car Accident, and Robbery—we aim to dissect the nuanced relationship between regret and human behavior. These scenarios, each with its unique context and set of outcomes, serve as fertile ground for investigating the dynamics of regret in the face of new experiences, luck, and misfortune.

Influential work by Kahneman and Miller in norm theory has highlighted the connection between abnormal behavior, heightened counterfactual thinking, and intensified feelings of regret. Their discussion of the concept of normality included experimental research illustrating various factors influencing the perception and interpretation of normality.

However, these analyses do not describe the relationship between the three elements: new experiences, luck, and compensation. This article, therefore, focuses on emphasizing which of these elements more significantly affects how individuals measure the extent of their regret,and then analyze the role of various factors from a psychological perspective.

The three questions below to analyze the three parts and the included experiments—Hitchhiker scenario, Car accident scenario, and Robbery scenario. The questions include: 
1. Why does attempting new experiences and encountering disasters often lead to greater regret? 
2. What is the relationship between luck as a measure of regret and the extent of regret
3: If we could measure regret, what is the relationship betweent compensation and regret??

The first question we address probes the paradoxical phenomenon where attempting new experiences and encountering subsequent disasters often lead to greater regret than sticking to familiar paths. This inquiry sheds light on the psychological mechanisms that amplify regret in the wake of novelty and adversity, suggesting that the allure of potential gains from uncharted territories comes with the shadow of heightened emotional risk. We explore the intricate dance between the desire for novelty and the aversion to negative outcomes, seeking to understand why our ventures into the unknown are so frequently shadowed by the specter of regret.

Secondly, we explore the relationship between luck, or the perceived randomness of outcomes, and the extent of regret experienced by individuals. This aspect of our analysis delves into the role of external factors beyond one's control in modulating the intensity of regret. By examining how individuals attribute outcomes to luck or personal agency, we uncover insights into the cognitive processes that influence the perception and experience of regret.

Lastly, our third question centers on the quantification of human regret, particularly in the context of new activities versus misfortune. Through Experiment 3, we seek to estimate the compensation required to mitigate the psychological trauma associated with regret stemming from engaging in new activities compared to that arising from unfortunate events. This investigation not only highlights the tangible impact of regret on well-being but also offers a glimpse into the potential for psychological interventions to alleviate the burden of this powerful emotion.

Together, these questions and scenarios weave a comprehensive narrative around the nature of regret, its triggers, and its effects on human behavior and psychological well-being. We can find that the connections among the three elements allow us to better assess the extent of people's regret.

we use R [@citeR] for all data wrangling and analysis and R packages tidyverse [@tidy], ggthemes [@ggthemes], ggprism [@ggprism] and patchwork [@patchwork] to produce the figures, kableExtra [@kableExtra] to produce the tables.

# Data
The data we used in this paper are obtained from the paper titled “The impact of past behavior normality on regret: replication and extension of three experiments of the exceptionality effect” (Kutscher & Feldman 2022). From the data in the article's tables, for Experiment 1,with Mr. Smith at 7.9% and Mr. Jones at a much higher 92.1%. A similar pattern emerges for injunctive social norms, where Mr. Smith scores 4.7%, contrasting sharply with Mr. Jones at 95.3%. Descriptive social norms show an obvious preference for routine behavior, as reflected in Mr. Smith's 90.6%, while Mr. Jones lags behind at 9.4%. Negative affect follows a parallel trend, with Mr. Smith at 7.3% and Mr. Jones at 92.7%. From Experiment 2 data, 81% of 277 respondents believed that Mr. White, who had recently switched to a new route home and then had an accident, would experience greater regret. Regarding luck, Mr. White (the exception) was seen as less fortunate compared to Mr. Adams (the routine), with 33% (114 participants) considering Mr. Adams as less lucky, while 66% (228 participants) viewed Mr. White as the less fortunate one.

## Proportion for perceived regret, injunctive social norms, descriptive social norms, negative affect and luck。

Regret measures the level of regret experienced by Mr. Smith (who frequently hitchhikes and was robbed today) and Mr. Jones (who almost never hitchhikes but was robbed on the one occasion he did today). Injunctive social norms represent the socially approved and accepted behaviors, while descriptive social norms represent criticized and socially disapproved behaviors. Finally, negative affect refers to the negative emotions felt by both individuals.

## Proportion for luck

The "luck variable" is from Experiment 2, where Mr. Adams was involved in an accident while driving home after work on his regular route. Mr. White experienced a similar accident while driving on a route he only takes when he desires a change of scenery. "Luck" represents an assessment of their perceived level of luck.

## Victim compensation and regret (with exception conditions combined)
The last figure titled "Victim Compensation and Regret (with Exception Conditions Combined)," presents an overview of two experiments, each with routine and exception conditions. The top two plots correspond to Experiment 1, while the bottom two plots represent Experiment 2, with error bars indicating standard errors. The scales range from 1 to 5 for compensation and 1 to 10 for regret. In Experiment 1, the routine condition includes 114 participants, with standard deviations of 2.37 for compensation and 1.12 for regret. The self-produced exception condition involves 115 participants, with standard deviations of 2.46 for compensation and 0.80 for regret. For the other-produced exception condition (113 participants), standard deviations are 2.76 for compensation and 0.94 for regret. Combining all exception conditions (228 participants) gets standard deviations of 2.61 for compensation and 0.87 for regret. In Experiment 2, routine conditions involve 116 participants, with standard deviations of 3.40 for compensation and 1.19 for regret. The self-produced exception condition (110 participants) shows standard deviations of 3.50 for compensation and 0.67 for regret. The other-produced exception condition (116 participants) has standard deviations of 3.66 for compensation and 0.98 for regret. Combining all exception conditions (226 participants) results in standard deviations of 3.57 for compensation and 0.86 for regret. These detailed data provide an understanding of victim compensation and regret across various experimental conditions.

# Results

## Overall evaluation in regre depending hitchhiker scenario

Figure \@ref(fig:fig1) is a replication of Figure 1 in the original paper, In this section, participants evaluated a classic scenario known as the hitchhiker scenario, to assess the impact of past behavior on the perception of regret following negative outcomes. 92% of the 315 participants believed that Mr. Jones, who almost never gives rides to others, felt a stronger sense of regret. The majority of participants believed that Mr. Jones, who almost never picks up hitchhikers, exhibits behavior that is more common in society (95% out of 326 respondents), whereas Mr. Smith, who frequently gives rides to hitchhikers, is more likely to be criticized by society (91% out of 310 respondents). For the assessment of negative affect, 93% of people believed that Mr. Jones experienced more negative affect compared to Mr. Smith. Overall, these two experiments provided very close replications, similar to the the original study.

```{r fig1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Rates for Regret, Social norms and Negative affect"}
data1 <- read.csv("/cloud/project/inputs/data/Ex1.csv", header = TRUE, stringsAsFactors = FALSE)

# Figure 1 replication

library(ggplot2)
library(reshape2)
library(ggplot2)
library(readr)
library(dplyr)

data <- data.frame(
  Category = c('Regret', 'Social norms (injunctive)', 'Social norms (descriptive)', 'Negative affect'),
  Smith_Routine = c(7.9, 4.7, 90.6, 7.3),
  Jones_Exception = c(92.1, 95.3, 9.4, 92.7)
)

long_data <- melt(data, id.vars = 'Category', variable.name = 'Group', value.name = 'Percentage')

fig1 <- ggplot(long_data, aes(x = Category, y = Percentage, fill = Group)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  scale_fill_grey(start = 0.8, end = 0.2) + # Adjust the grey scale to match the example
  theme_minimal() +
  labs(
    fill = "", 
    x = "",    
    y = ""     
  ) +
  theme(
    legend.position = "bottom", 
    axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels if necessary
    plot.title = element_text(hjust = 0.5) # Center the plot title if you add one
  )
fig1
```

## Car accident scenario

Figure 2 is a replication of Figure 2 in the original paper, and shows the Proportions for perceived regret and luck.This graph compares the relationship between regret and luck, with 81% of 277 respondents believing that Mr. White, who had just changed to a new route home and experienced an accident, would feel more regretful.As for luck, Mr. White (exception) was considered less lucky than Mr. Adams (routine), with 114 participants (33%) viewing Mr. Adams as less lucky, whereas 228 (66%) perceived Mr. White to be less lucky.

```{r fig2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Proportions for perceived regret and luck"}

library(ggplot2)
library(reshape2)

df <- read.csv("/cloud/project/inputs/data/table2.csv", header = TRUE, stringsAsFactors = FALSE)

df_clean <- df[, -c(2, 4)]
names(df_clean) <- c("Category", "Regret", "Luck")

df_melted <- melt(df_clean, id.vars = "Category", variable.name = "Outcome", value.name = "Percentage")

fig2 <- ggplot(df_melted, aes(x = Outcome, y = Percentage, fill = Category)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
  scale_fill_manual(values=c("Routine Adams" = "black", "Exception White" = "grey")) +
  labs(x = "", y = "Percentage", fill = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")

fig2
```

Figure 3 In the first graph, we can observe two distributions corresponding to the compensation under routine behavior conditions and exception conditions. The compensation amounts are represented on a scale from 0 to 10, which might indicate a range from $0 to $1,000,000. Under the routine behavior condition, the data distribution appears to be more concentrated, suggesting that most participants' compensation amounts were relatively close to each other, and the median seems to be near the middle value of the scale. Under the exception condition, the distribution of compensation data is broader, indicating a greater variance in the participants' assessments of the compensation amount.

The second graph illustrating the distribution of reported regret under routine and exception conditions.In the routine condition, the distribution of regret is narrower and more concentrated around the lower end of the scale, suggesting that participants generally felt less regret for the routine situation. The median regret rating, indicated by the horizontal line inside the violin plot, appears to be below the midpoint of the scale.In the exception condition, the regret distribution is wider at the top, indicating that there is a greater variance in how much regret participants felt, with more reports of higher levels of regret. The median regret rating in this condition is higher than in the routine condition, as seen by its position above the midpoint of the scale.

```{r fig3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="compensation and reported regret "}

data <- read.csv("/cloud/project/inputs/data/osf-past-normality-regret-replication-exp1-data.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
data$Sc1_regret <- factor(data$Sc1_regret, levels = c(1,2), ordered = TRUE, labels = c("Exception Jones", "Routine Smith"))
data$sc1_socnorms1 <- factor(data$sc1_socnorms1, levels = c(1,2), ordered = TRUE, labels = c("Exception Jones", "Routine Smith"))
data$sc1_socnorms2 <- factor(data$sc1_socnorms2, levels = c(1,2), ordered = TRUE, labels = c("Exception Jones", "Routine Smith"))
data$sc1_combinednorms <- factor(data$sc1_combinednorms, levels = c(1,2), ordered = TRUE, labels = c("Exception Jones", "Routine Smith"))

data$Sc3condition1 <- NA
data$compensationagg1 <- NA
data$regretagg1 <- NA

data <- transform(data, Sc3condition1 = ifelse(!is.na(Sc3_C1_text), 1, ifelse(!is.na(Sc3_C2_text) | !is.na(Sc3_C3_text), 2, NA)),
                  compensationagg1 = ifelse(!is.na(Sc3_C1_text), sc3_c1_compensation, ifelse(!is.na(Sc3_C2_text), sc3_c2_compensation, ifelse(!is.na(Sc3_C3_text), sc3_c3_compensation, NA))),
                  regretagg1 = ifelse(!is.na(Sc3_C1_text), sc3_c1_regret, ifelse(!is.na(Sc3_C2_text), sc3_c2_regret, ifelse(!is.na(Sc3_C3_text), sc3_c3_regret, NA))))

data$Sc3conditionl1 <- factor(data$Sc3condition1, levels = c(1, 2), labels = c("Routine", "Exception"))

data$compensationaggrecoded1 <- data$compensationagg1 - 1
data$regretaggrecoded1 <- data$regretagg1 + 1

ggplot(data, aes(x = Sc3conditionl1, y = compensationaggrecoded1)) + 
  geom_violin(trim = FALSE) + 
  geom_jitter(shape = 16, position = position_jitter(height = 0.1, width = 0.2)) + 
  theme_apa() +
  labs(x = "Condition", y = "Compensation") +
  scale_y_continuous(minor_breaks = seq(-10, 15, 5)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = position_dodge(width = 0.9), width = 0.2) +
  stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3)

ggplot(data, aes(x = Sc3conditionl1, y = regretaggrecoded1)) + 
  geom_violin(trim = FALSE) + 
  geom_jitter(shape = 16, position = position_jitter(height = 0.1, width = 0.2)) + 
  theme_apa() +
  labs(x = "Condition", y = "Regret") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = position_dodge(width = 0.9), width = 0.2) +
  stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3)

```

# Discussion

## 1. Why does attempting new experiences and encountering disasters often lead to greater regret?

To further explore this concept, we first check for figure1, it's apparent that in the survey, more people believe Mr. Jones, who rarely offers rides, would feel a stronger sense of regret for being robbed due to offering a ride just once. Firstly, the habit of offering rides initially influences one's psychological expectations. For instance, someone who has never offered rides before would tend to believe that their kind act, or an exception, would to some extent help others or receive positive psychological feedback. However, conversely, when their psychological expectations are not met, the sense of psychological disparity comes into play, leading people to feel that acts of kindness might instead invite unnecessary trouble, thereby intensifying their psychological fixation. Positive feedback would make them happier, while negative impacts would bring about more feelings of regret.

From a psychological perspective，counterfactual thinking involves individuals imagining how different actions could have led to different outcomes, particularly after experiencing adverse results. This mental process becomes especially pronounced when attempting new experiences and encountering disasters, as people tend to evaluate their decisions by constructing scenarios that differ from what actually happened. For instance, consider someone who decides to engage in a high-risk investment, hoping for significant returns, but instead faces substantial financial losses. In this context, counterfactual thinking might manifest as thoughts like, "If I hadn't invested in that project, my financial situation would be better now." Such thinking not only deepens the individual's sense of regret but might also lead to doubts about their decision-making abilities. Given that individuals tend to assess the positive outcomes of unchosen alternatives, this mode of thinking amplifies regret over the actual choice made. Therefore, counterfactual thinking reveals how people enhance their feelings of regret by comparing actual outcomes with imagined, potentially better ones, especially when venturing into new experiences that yield unexpected negative consequences.

we secondly need to verify whether accidents in exceptional behaviors lead to more regret compared to accidents in routine behaviors. Venturing into new experiences and facing setbacks often leads to heightened feelings of regret, a phenomenon deeply rooted in our psychological responses to deviations from routine. When individuals engage in behaviors that stray from their norm—deciding on a whim to take a different route home, trying out an unconventional restaurant, or making an impulsive purchase—any negative outcomes tend to be felt more acutely. This intensified regret is linked to the concept of "exceptionality," where actions that stand out as exceptions to regular patterns are more likely to trigger stronger counterfactual thinking. Essentially, when outcomes don't go as planned, our minds more easily imagine alternative scenarios where sticking to the routine could have avoided the undesired result.

## 2. What is the relationship between luck as a measure of regret and the extent of regret

From the article, we can see that among 277 people, 81% believe that a person named White who was involved in a similar accident while driving on a route he only took for a change of scenery would feel greater regret and worse luck. Therefore, the factor of luck often has a direct relationship with the level of regret. When someone encounters a random event, the worse their luck, the stronger their feeling of regret. For example, Mr. Adams had an accident while driving home on his usual route after work. Mr. White had a similar accident while he was driving on a route he only took when he wanted a change of scenery. In such cases, for accidents that happen under normal circumstances, people might think that it was something bound to happen, believing that no matter what, if they were on that road, they would have faced the same situation, and thus, they wouldn't particularly feel unlucky or have much regret. However, people tend to attribute accidents that happen under unusual circumstances to bad luck. Mr. White's temporary choice led to a bad outcome. Since there was no way to weigh the probabilities of different choices at the moment of decision, people are more likely to attribute the event to bad luck, resulting in a stronger sense of regret.

The way people understand luck in relation to regret is detailed and subtle. Essentially, luck is about the element of chance in what happens after we make decisions or take actions. When people think that what happens is down to luck, their feelings of regret can change depending on how much control they think they had over the situation. If someone believes a bad outcome was just unlucky, they might not feel as much regret because they see it as something they couldn't avoid. On the other hand, if they see the outcome as the result of their own choices, the regret can be stronger because they think they could have chosen differently to prevent the bad result. This shows that how much regret people feel isn't just about what happens, but also how they interpret these happenings in terms of luck and their own actions.

## 3: If we could measure regret, what is the relationship betweent compensation and regret?

To comprehensively measure the degree of regret, we combined studies one and two to address this issue. Imagine a person, White, who usually visits a familiar store but encounters a robbery there, losing an arm in the struggle. Another person, Smith, is forced to go to a new store under unusual circumstances and faces the same tragedy. According to the question, people generally believe the amount of money needed to compensate for their losses ranges from $0 to $1,000,000. From the density of points in the article's graph, For White, the chart shows points densely clustered around the midpoint of $500,000, with few believing that compensation should exceed this figure. In contrast, the level of regret for Smith is stronger, resulting in a more dispersed distribution of compensation, with most people believing that compensation should be between ¥500,000 and ¥1,000,000. This suggests that the degree of regret is directly proportional to the compensation expected.

The interplay between counterfactual thinking and the consequent feelings of regret significantly influences individuals' expectations for compensation, a dynamic vividly illustrated through the experiences of White and Smith. Counterfactual thinking, the cognitive process of envisioning alternative scenarios that did not occur, can intensify feelings of regret, especially when these imagined outcomes are perceived as more favorable than the actual events. For White, who frequented a familiar store and experienced a robbery resulting in severe injury, the compensation expectations are densely clustered around $500,000, suggesting that the routine nature of his visit might have led to less intense counterfactual thinking and subsequently, a more contained range of expected compensation. In contrast, Smith, who ventured into an unfamiliar store under unusual circumstances and encountered the same fate, experiences stronger regret, fueled by more potent counterfactual thoughts such as "If only I had avoided making that unusual choice." This heightened level of regret is reflected in the more dispersed compensation expectations, with most believing that the figure should lie between ¥500,000 and ¥1,000,000. This scenario underlines how the depth and focus of counterfactual thinking can dramatically shape an individual's perception of their loss and their valuation of necessary compensation. It showcases the complex relationship between cognitive processes and emotional responses, demonstrating that individuals' compensation expectations are not only a reflection of their physical losses but also an attempt to mitigate the emotional distress caused by their reflections on what could have been, thus highlighting the profound impact of counterfactual thinking on human behavior and expectations in the aftermath of traumatic events.

# Conclusion

This article delves into how people perceive and process regret across various situations, from offering help that leads to negative outcomes, to the role luck plays in our feelings of regret, and finally, how we value compensation after unfortunate events. It highlights how our thoughts about "what could have been," our views on chance, and our expectations for making amends are deeply intertwined with our emotional reactions. The study reveals the complexity of our emotional world and how our minds influence our feelings and choices after difficult situations. By understanding these relationships, we gain insight into the complex nature of human emotions and the significant impact our cognitive and emotional responses have on how we navigate and make sense of challenges in our lives.

## Weaknesses and next steps

Due to the degree of regret, luck, and negative impacts being largely related to emotional value, it's challenging to measure their monetary value using tangible factors. Therefore, the variable of compensation is quite inaccurate, being only derived from survey data without actual reference value. For the next step, we are inclined to identify more suitable variables to quantify the sense of regret, thereby enabling a better assessment of the compensation issue.

\newpage

# References

### 2024-02-15 outputs/paper/Paper_2.qmd
1: organize all files, include delect useless files 2: update read me files, add a LLM introduction

@@
 
-Figure \@ref(fig:fig1) is a replication of Figure 1 in the original paper, In this section, participants evaluated a classic scenario known as the hitchhiker scenario, to assess the impact of past behavior on the perception of regret following negative outcomes.
+Figure1 is a replication of Figure 1 in the original paper, In this section, participants evaluated a classic scenario known as the hitchhiker scenario, to assess the impact of past behavior on the perception of regret following negative outcomes.
 92% of the 315 participants believed that Mr.
@@
     axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels if necessary
-    plot.title = element_text(hjust = 0.5) # Center the plot title if you add one
+    plot.title = element_text(hjust = 0.5) 
   )
@@
 
-ggplot(data, aes(x = Sc3conditionl1, y = compensationaggrecoded1)) + 
+fig3 <- ggplot(data, aes(x = Sc3conditionl1, y = compensationaggrecoded1)) + 
   geom_violin(trim = FALSE) + 
@@
 
-ggplot(data, aes(x = Sc3conditionl1, y = regretaggrecoded1)) + 
+fig4 <- ggplot(data, aes(x = Sc3conditionl1, y = regretaggrecoded1)) + 
   geom_violin(trim = FALSE) + 
@@
 
+fig3 
+fig4
 ```

### 2024-02-15 outputs/paper/Paper_2.qmd
update

@@
 This emotional response, characterized by a sense of loss or disappointment over what might have been, emerges from the recognition that different decisions or circumstances could have led to more favorable outcomes.
-While regret is a universal experience, its intensity and impact can vary widely among individuals and situations.
+While regret is a universal experience, its intensity and impact can vary widely among individuals and situations (Byre,2016;Roese&·Epstude,2017;Kahneman&Tversky,1982).
 In this article, we delve into the complex nature of regret, exploring its triggers, consequences, and the psychological underpinnings that shape this emotional landscape.
@@
 Influential work by Kahneman and Miller in norm theory has highlighted the connection between abnormal behavior, heightened counterfactual thinking, and intensified feelings of regret.
-Their discussion of the concept of normality included experimental research illustrating various factors influencing the perception and interpretation of normality.
+Their discussion of the concept of normality included experimental research illustrating various factors influencing the perception and interpretation of normality (Feldman&Albaracin,2017;Feldman,2018).
 
@@
 The first question we address probes the paradoxical phenomenon where attempting new experiences and encountering subsequent disasters often lead to greater regret than sticking to familiar paths.
-This inquiry sheds light on the psychological mechanisms that amplify regret in the wake of novelty and adversity, suggesting that the allure of potential gains from uncharted territories comes with the shadow of heightened emotional risk.
+This inquiry sheds light on the psychological mechanisms that amplify regret in the wake of novelty and adversity, suggesting that the allure of potential gains from uncharted territories comes with the shadow of heightened emotional risk (Kahneman & Miller 1986).
 We explore the intricate dance between the desire for novelty and the aversion to negative outcomes, seeking to understand why our ventures into the unknown are so frequently shadowed by the specter of regret.
@@
 Lastly, our third question centers on the quantification of human regret, particularly in the context of new activities versus misfortune.
-Through Experiment 3, we seek to estimate the compensation required to mitigate the psychological trauma associated with regret stemming from engaging in new activities compared to that arising from unfortunate events.
+Through Experiment 3, we seek to estimate the compensation required to mitigate the psychological trauma associated with regret stemming from engaging in new activities compared to that arising from unfortunate events (Kahneman & Miller 1986).
 This investigation not only highlights the tangible impact of regret on well-being but also offers a glimpse into the potential for psychological interventions to alleviate the burden of this powerful emotion.
@@
 Together, these questions and scenarios weave a comprehensive narrative around the nature of regret, its triggers, and its effects on human behavior and psychological well-being.
-We can find that the connections among the three elements allow us to better assess the extent of people's regret.
+We can find that the connections among the three elements allow us to better assess the extent of people's regret (Kahneman & Miller 1986).
 

### 2024-02-15 outputs/paper/Paper_2.qmd
update

@@
 However, conversely, when their psychological expectations are not met, the sense of psychological disparity comes into play, leading people to feel that acts of kindness might instead invite unnecessary trouble, thereby intensifying their psychological fixation.
-Positive feedback would make them happier, while negative impacts would bring about more feelings of regret.
+Positive feedback would make them happier, while negative impacts would bring about more feelings of regret (Kutscher & Feldman 2019).
 
 From a psychological perspective，counterfactual thinking involves individuals imagining how different actions could have led to different outcomes, particularly after experiencing adverse results.
-This mental process becomes especially pronounced when attempting new experiences and encountering disasters, as people tend to evaluate their decisions by constructing scenarios that differ from what actually happened.
+This mental process becomes especially pronounced when attempting new experiences and encountering disasters, as people tend to evaluate their decisions by constructing scenarios that differ from what actually happened (Roese & Olson, 1997).
 For instance, consider someone who decides to engage in a high-risk investment, hoping for significant returns, but instead faces substantial financial losses.
@@
 Given that individuals tend to assess the positive outcomes of unchosen alternatives, this mode of thinking amplifies regret over the actual choice made.
-Therefore, counterfactual thinking reveals how people enhance their feelings of regret by comparing actual outcomes with imagined, potentially better ones, especially when venturing into new experiences that yield unexpected negative consequences.
+Therefore, counterfactual thinking reveals how people enhance their feelings of regret by comparing actual outcomes with imagined, potentially better ones, especially when venturing into new experiences that yield unexpected negative consequences (Roese & Olson, 1997).
 
@@
 When individuals engage in behaviors that stray from their norm—deciding on a whim to take a different route home, trying out an unconventional restaurant, or making an impulsive purchase—any negative outcomes tend to be felt more acutely.
-This intensified regret is linked to the concept of "exceptionality," where actions that stand out as exceptions to regular patterns are more likely to trigger stronger counterfactual thinking.
+This intensified regret is linked to the concept of "exceptionality," where actions that stand out as exceptions to regular patterns are more likely to trigger stronger counterfactual thinking (Briazu et al, 2017).
 Essentially, when outcomes don't go as planned, our minds more easily imagine alternative scenarios where sticking to the routine could have avoided the undesired result.
@@
 White had a similar accident while he was driving on a route he only took when he wanted a change of scenery.
-In such cases, for accidents that happen under normal circumstances, people might think that it was something bound to happen, believing that no matter what, if they were on that road, they would have faced the same situation, and thus, they wouldn't particularly feel unlucky or have much regret.
+In such cases, for accidents that happen under normal circumstances, people might think that it was something bound to happen, believing that no matter what, if they were on that road, they would have faced the same situation, and thus, they wouldn't particularly feel unlucky or have much regret (Kutscher & Feldman 2019).
 However, people tend to attribute accidents that happen under unusual circumstances to bad luck.
@@
 If someone believes a bad outcome was just unlucky, they might not feel as much regret because they see it as something they couldn't avoid.
-On the other hand, if they see the outcome as the result of their own choices, the regret can be stronger because they think they could have chosen differently to prevent the bad result.
+On the other hand, if they see the outcome as the result of their own choices, the regret can be stronger because they think they could have chosen differently to prevent the bad result (Kutscher & Feldman 2019).
 This shows that how much regret people feel isn't just about what happens, but also how they interpret these happenings in terms of luck and their own actions.
@@
 This scenario underlines how the depth and focus of counterfactual thinking can dramatically shape an individual's perception of their loss and their valuation of necessary compensation.
-It showcases the complex relationship between cognitive processes and emotional responses, demonstrating that individuals' compensation expectations are not only a reflection of their physical losses but also an attempt to mitigate the emotional distress caused by their reflections on what could have been, thus highlighting the profound impact of counterfactual thinking on human behavior and expectations in the aftermath of traumatic events.
+It showcases the complex relationship between cognitive processes and emotional responses, demonstrating that individuals' compensation expectations are not only a reflection of their physical losses but also an attempt to mitigate the emotional distress caused by their reflections on what could have been, thus highlighting the profound impact of counterfactual thinking on human behavior and expectations in the aftermath of traumatic events (Kutscher & Feldman 2019).
 

### 2024-02-15 outputs/paper/Paper_2.qmd
update

@@
 White had a similar accident while he was driving on a route he only took when he wanted a change of scenery.
-In such cases, for accidents that happen under normal circumstances, people might think that it was something bound to happen, believing that no matter what, if they were on that road, they would have faced the same situation, and thus, they wouldn't particularly feel unlucky or have much regret (Kutscher & Feldman 2019).
+In such cases, for accidents that happen under normal circumstances, people might think that it was something bound to happen, believing that no matter what, if they were on that road, they would have faced the same situation, and thus, they wouldn't particularly feel unlucky or have much regret (Kutscher & Feldman, 2019).
 However, people tend to attribute accidents that happen under unusual circumstances to bad luck.
@@
 If someone believes a bad outcome was just unlucky, they might not feel as much regret because they see it as something they couldn't avoid.
-On the other hand, if they see the outcome as the result of their own choices, the regret can be stronger because they think they could have chosen differently to prevent the bad result (Kutscher & Feldman 2019).
+On the other hand, if they see the outcome as the result of their own choices, the regret can be stronger because they think they could have chosen differently to prevent the bad result (Kutscher & Feldman, 2019).
 This shows that how much regret people feel isn't just about what happens, but also how they interpret these happenings in terms of luck and their own actions.
@@
 This scenario underlines how the depth and focus of counterfactual thinking can dramatically shape an individual's perception of their loss and their valuation of necessary compensation.
-It showcases the complex relationship between cognitive processes and emotional responses, demonstrating that individuals' compensation expectations are not only a reflection of their physical losses but also an attempt to mitigate the emotional distress caused by their reflections on what could have been, thus highlighting the profound impact of counterfactual thinking on human behavior and expectations in the aftermath of traumatic events (Kutscher & Feldman 2019).
+It showcases the complex relationship between cognitive processes and emotional responses, demonstrating that individuals' compensation expectations are not only a reflection of their physical losses but also an attempt to mitigate the emotional distress caused by their reflections on what could have been, thus highlighting the profound impact of counterfactual thinking on human behavior and expectations in the aftermath of traumatic events (Kutscher & Feldman, 2019).
 
 # Conclusion
-
 This article delves into how people perceive and process regret across various situations, from offering help that leads to negative outcomes, to the role luck plays in our feelings of regret, and finally, how we value compensation after unfortunate events.

### 2024-02-15 outputs/paper/Paper_2.qmd
update

@@
 Counterfactual thinking, the cognitive process of envisioning alternative scenarios that did not occur, can intensify feelings of regret, especially when these imagined outcomes are perceived as more favorable than the actual events.
-For White, who frequented a familiar store and experienced a robbery resulting in severe injury, the compensation expectations are densely clustered around $500,000, suggesting that the routine nature of his visit might have led to less intense counterfactual thinking and subsequently, a more contained range of expected compensation.
+For White, who frequented a familiar store and experienced a robbery resulting in severe injury, the compensation expectations are densely clustered around $500,000, suggesting that the routine nature of his visit might have led to less intense counterfactual thinking and subsequently, a more contained range of expected compensation (Kutscher & Feldman 2019).
 In contrast, Smith, who ventured into an unfamiliar store under unusual circumstances and encountered the same fate, experiences stronger regret, fueled by more potent counterfactual thoughts such as "If only I had avoided making that unusual choice." This heightened level of regret is reflected in the more dispersed compensation expectations, with most believing that the figure should lie between ¥500,000 and ¥1,000,000.

### 2024-02-15 outputs/paper/Paper_2.qmd
1: add more explaination for data parts 2: add a LLM file to introduce the LLM useage 3: rename the PDF file

--- 2024-02-15 outputs/paper/Paper_2.qmd
@@
 This emotional response, characterized by a sense of loss or disappointment over what might have been, emerges from the recognition that different decisions or circumstances could have led to more favorable outcomes.
-While regret is a universal experience, its intensity and impact can vary widely among individuals and situations (Byre,2016;Roese&·Epstude,2017;Kahneman&Tversky,1982).
+While regret is a universal experience, its intensity and impact can vary widely among individuals and situations (@Byrne2016).
 In this article, we delve into the complex nature of regret, exploring its triggers, consequences, and the psychological underpinnings that shape this emotional landscape.
@@
 Influential work by Kahneman and Miller in norm theory has highlighted the connection between abnormal behavior, heightened counterfactual thinking, and intensified feelings of regret.
-Their discussion of the concept of normality included experimental research illustrating various factors influencing the perception and interpretation of normality (Feldman&Albaracin,2017;Feldman,2018).
+(@Miller)Their discussion of the concept of normality included experimental research illustrating various factors influencing the perception and interpretation of normality (@Feldman2018).
 
@@
 The first question we address probes the paradoxical phenomenon where attempting new experiences and encountering subsequent disasters often lead to greater regret than sticking to familiar paths.
-This inquiry sheds light on the psychological mechanisms that amplify regret in the wake of novelty and adversity, suggesting that the allure of potential gains from uncharted territories comes with the shadow of heightened emotional risk (Kahneman & Miller 1986).
+This inquiry sheds light on the psychological mechanisms that amplify regret in the wake of novelty and adversity, suggesting that the allure of potential gains from uncharted territories comes with the shadow of heightened emotional risk (@Kahneman).
 We explore the intricate dance between the desire for novelty and the aversion to negative outcomes, seeking to understand why our ventures into the unknown are so frequently shadowed by the specter of regret.
@@
 Lastly, our third question centers on the quantification of human regret, particularly in the context of new activities versus misfortune.
-Through Experiment 3, we seek to estimate the compensation required to mitigate the psychological trauma associated with regret stemming from engaging in new activities compared to that arising from unfortunate events (Kahneman & Miller 1986).
+Through Experiment 3, we seek to estimate the compensation required to mitigate the psychological trauma associated with regret stemming from engaging in new activities compared to that arising from unfortunate events (@kahneman1986norm).
 This investigation not only highlights the tangible impact of regret on well-being but also offers a glimpse into the potential for psychological interventions to alleviate the burden of this powerful emotion.
@@
 Together, these questions and scenarios weave a comprehensive narrative around the nature of regret, its triggers, and its effects on human behavior and psychological well-being.
-We can find that the connections among the three elements allow us to better assess the extent of people's regret (Kahneman & Miller 1986).
+We can find that the connections among the three elements allow us to better assess the extent of people's regret (@kahneman1986norm).
 
@@
 # Data
-The data we used in this paper are obtained from the paper titled “The impact of past behavior normality on regret: replication and extension of three experiments of the exceptionality effect” (Kutscher & Feldman 2022).
+The data we used in this paper are obtained from the paper titled “The impact of past behavior normality on regret: replication and extension of three experiments of the exceptionality effect” (@Roese).
+All the data below were collected from the tables and data provided under the original article "The impact of past behaviour normality on regret: replication and extension of three experiments of the exceptionality effect.
 From the data in the article's tables, for Experiment 1,with Mr.
@@
 White as the less fortunate one.
+In part3, the density of data points explains the majority's view on how much compensation is deemed appropriate in relation to their internal sense of regret.
+The first graph indicates that most people feel that they should receive a compensation of 500,000, whereas the distribution of data points in the second is more dispersed, with the majority feeling that they should receive more, ranging from 500,000 to 1,000,000.
+The second graph is similar, only the variable has been changed to regret, with more people feeling that the second individual has greater regret.
 
@@
 
-data <- data.frame(
-  Category = c('Regret', 'Social norms (injunctive)', 'Social norms (descriptive)', 'Negative affect'),
-  Smith_Routine = c(7.9, 4.7, 90.6, 7.3),
-  Jones_Exception = c(92.1, 95.3, 9.4, 92.7)
-)
+library(ggplot2)
+library(reshape2)
 
-long_data <- melt(data, id.vars = 'Category', variable.name = 'Group', value.name = 'Percentage')
+# Read the data from Exp1.csv
+data <- read.csv("/cloud/project/inputs/data/Ex1.csv")
+data <- data[-1,]
 
-fig1 <- ggplot(long_data, aes(x = Category, y = Percentage, fill = Group)) +
+# Select columns by index assuming the order is as follows:
+# Condition name, Percentage for Regret, Percentage for Social norm (injunctive),
+# Percentage for Social norms (descriptive), Percentage for Negative affect
+data <- data[, c(1, 3, 5, 7, 9)]
+
+# Assigning the correct names to the columns
+names(data) <- c("Condition", "Regret", "Social_norm_injunctive", "Social_norms_descriptive", "Negative_affect")
+
+# Convert the percentage columns to numeric (as.character is used to handle factors)
+data$Regret <- as.numeric(as.character(data$Regret))
+data$Social_norm_injunctive <- as.numeric(as.character(data$Social_norm_injunctive))
+data$Social_norms_descriptive <- as.numeric(as.character(data$Social_norms_descriptive))
+data$Negative_affect <- as.numeric(as.character(data$Negative_affect))
+
+# Melt the data for use with ggplot2
+data_melted <- melt(data, id.vars = 'Condition')
+
+# Plotting
+ggplot(data_melted, aes(x = variable, y = value, fill = Condition)) + 
   geom_bar(stat = 'identity', position = 'dodge') +
-  scale_fill_grey(start = 0.8, end = 0.2) + # Adjust the grey scale to match the example
+  scale_fill_grey(start = 0.8, end = 0.2) +  # Adjust the colors to match the image
   theme_minimal() +
-  labs(
-    fill = "", 
-    x = "",    
-    y = ""     
-  ) +
-  theme(
-    legend.position = "bottom", 
-    axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels if necessary
-    plot.title = element_text(hjust = 0.5) 
-  )
-fig1
+  labs(x = '', y = '') +  # Removing axis labels as in the image
+  theme(legend.title = element_blank(),  # Removing legend title
+        legend.position = 'bottom') +  # Moving legend to bottom
+  scale_y_continuous(labels = scales::percent_format())
 ```
@@
 
-df <- read.csv("/cloud/project/inputs/data/table2.csv", header = TRUE, stringsAsFactors = FALSE)
+library(ggplot2)
+library(reshape2)
 
-df_clean <- df[, -c(2, 4)]
-names(df_clean) <- c("Category", "Regret", "Luck")
+data <- read.csv("/cloud/project/inputs/data/table2.csv")
 
-df_melted <- melt(df_clean, id.vars = "Category", variable.name = "Outcome", value.name = "Percentage")
+data <- data[-1,]
+data <- data[, c(1, 3, 5)]
 
-fig2 <- ggplot(df_melted, aes(x = Outcome, y = Percentage, fill = Category)) +
-  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
-  scale_fill_manual(values=c("Routine Adams" = "black", "Exception White" = "grey")) +
-  labs(x = "", y = "Percentage", fill = "") +
+names(data) <- c("Condition", "Regret", "Luck")
+data$Regret <- as.numeric(as.character(data$Regret))
+data$Luck <- as.numeric(as.character(data$Luck))
+data_melted <- melt(data, id.vars = 'Condition')
+
+ggplot(data_melted, aes(x = variable, y = value, fill = Condition)) + 
+  geom_bar(stat = 'identity', position = 'dodge') +
+  scale_fill_grey(start = 0.8, end = 0.2) +  
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.position = "bottom")
-
-fig2
+  labs(x = '', y = '') +  # Removing axis labels as in the image
+  theme(legend.title = element_blank(),  
+        legend.position = 'bottom') +  
+  scale_y_continuous(labels = scales::percent_format()) 
 ```
@@
 However, conversely, when their psychological expectations are not met, the sense of psychological disparity comes into play, leading people to feel that acts of kindness might instead invite unnecessary trouble, thereby intensifying their psychological fixation.
-Positive feedback would make them happier, while negative impacts would bring about more feelings of regret (Kutscher & Feldman 2019).
+Positive feedback would make them happier, while negative impacts would bring about more feelings of regret (@Kutscher).
 
 From a psychological perspective，counterfactual thinking involves individuals imagining how different actions could have led to different outcomes, particularly after experiencing adverse results.
-This mental process becomes especially pronounced when attempting new experiences and encountering disasters, as people tend to evaluate their decisions by constructing scenarios that differ from what actually happened (Roese & Olson, 1997).
+This mental process becomes especially pronounced when attempting new experiences and encountering disasters, as people tend to evaluate their decisions by constructing scenarios that differ from what actually happened (@roese1997affective).
 For instance, consider someone who decides to engage in a high-risk investment, hoping for significant returns, but instead faces substantial financial losses.
@@
 Given that individuals tend to assess the positive outcomes of unchosen alternatives, this mode of thinking amplifies regret over the actual choice made.
-Therefore, counterfactual thinking reveals how people enhance their feelings of regret by comparing actual outcomes with imagined, potentially better ones, especially when venturing into new experiences that yield unexpected negative consequences (Roese & Olson, 1997).
+Therefore, counterfactual thinking reveals how people enhance their feelings of regret by comparing actual outcomes with imagined, potentially better ones, especially when venturing into new experiences that yield unexpected negative consequences (@roese1997counterfactual).
 
@@
 When individuals engage in behaviors that stray from their norm—deciding on a whim to take a different route home, trying out an unconventional restaurant, or making an impulsive purchase—any negative outcomes tend to be felt more acutely.
-This intensified regret is linked to the concept of "exceptionality," where actions that stand out as exceptions to regular patterns are more likely to trigger stronger counterfactual thinking (Briazu et al, 2017).
+This intensified regret is linked to the concept of "exceptionality," where actions that stand out as exceptions to regular patterns are more likely to trigger stronger counterfactual thinking (@briazu2017undoing).
 Essentially, when outcomes don't go as planned, our minds more easily imagine alternative scenarios where sticking to the routine could have avoided the undesired result.
@@
 White had a similar accident while he was driving on a route he only took when he wanted a change of scenery.
-In such cases, for accidents that happen under normal circumstances, people might think that it was something bound to happen, believing that no matter what, if they were on that road, they would have faced the same situation, and thus, they wouldn't particularly feel unlucky or have much regret (Kutscher & Feldman, 2019).
+In such cases, for accidents that happen under normal circumstances, people might think that it was something bound to happen, believing that no matter what, if they were on that road, they would have faced the same situation, and thus, they wouldn't particularly feel unlucky or have much regret (@Kutscher).
 However, people tend to attribute accidents that happen under unusual circumstances to bad luck.
@@
 If someone believes a bad outcome was just unlucky, they might not feel as much regret because they see it as something they couldn't avoid.
-On the other hand, if they see the outcome as the result of their own choices, the regret can be stronger because they think they could have chosen differently to prevent the bad result (Kutscher & Feldman, 2019).
+On the other hand, if they see the outcome as the result of their own choices, the regret can be stronger because they think they could have chosen differently to prevent the bad result (@Kutscher).
 This shows that how much regret people feel isn't just about what happens, but also how they interpret these happenings in terms of luck and their own actions.
@@
 Counterfactual thinking, the cognitive process of envisioning alternative scenarios that did not occur, can intensify feelings of regret, especially when these imagined outcomes are perceived as more favorable than the actual events.
-For White, who frequented a familiar store and experienced a robbery resulting in severe injury, the compensation expectations are densely clustered around $500,000, suggesting that the routine nature of his visit might have led to less intense counterfactual thinking and subsequently, a more contained range of expected compensation.
+For White, who frequented a familiar store and experienced a robbery resulting in severe injury, the compensation expectations are densely clustered around $500,000, suggesting that the routine nature of his visit might have led to less intense counterfactual thinking and subsequently, a more contained range of expected compensation (@Kutscher).
 In contrast, Smith, who ventured into an unfamiliar store under unusual circumstances and encountered the same fate, experiences stronger regret, fueled by more potent counterfactual thoughts such as "If only I had avoided making that unusual choice." This heightened level of regret is reflected in the more dispersed compensation expectations, with most believing that the figure should lie between ¥500,000 and ¥1,000,000.
 This scenario underlines how the depth and focus of counterfactual thinking can dramatically shape an individual's perception of their loss and their valuation of necessary compensation.
-It showcases the complex relationship between cognitive processes and emotional responses, demonstrating that individuals' compensation expectations are not only a reflection of their physical losses but also an attempt to mitigate the emotional distress caused by their reflections on what could have been, thus highlighting the profound impact of counterfactual thinking on human behavior and expectations in the aftermath of traumatic events (Kutscher & Feldman, 2019).
+It showcases the complex relationship between cognitive processes and emotional responses, demonstrating that individuals' compensation expectations are not only a reflection of their physical losses but also an attempt to mitigate the emotional distress caused by their reflections on what could have been, thus highlighting the profound impact of counterfactual thinking on human behavior and expectations in the aftermath of traumatic events (@Kutscher).
 

### 2024-02-15 replication/scripts/fig1_replication.R
1: add the figure 1, 2 and 3 replication to replication folder. 2: delect the useless replication files

@@ -1,88 +1,42 @@
-#### Preamble ####
-# Purpose: Figure 1 replication
-# Author: Kimlin Chin
-# Date: 22 February 2022
-# Contact: kimlin.chin@mail.utoronto.ca
-# License: MIT
+# Figure 1 replication
 
-## Data Sources:
-
-# Martin JA, Hamilton BE, Osterman MJK, et al. Births: Final data for 2015. 
-# National vital statistics report; vol 66, no 1. Hyattsville, MD: National Center for Health Statistics. 
-# 2017. https://www.cdc.gov/nchs/data/nvsr/nvsr66/nvsr66_01.pdf.
-# Data can be found on page 35,
-# Table 9. Birth rates, by live-birth order and race and Hispanic origin of mother: United States, 1980–2015
-# Columns: Fertility Rate, All races[years 1980-2015]
-
-# Martin JA, Hamilton BE, Osterman MJK, Driscoll AK. Births: Final Data for 2019. National
-# Vital Statistics Reports; vol 70 no 2. Hyattsville, MD: National Center for Health
-# Statistics. 2021. DOI: https://dx.doi.org/10.15620/cdc:100472.
-# Data can be found on page 13,
-# Table 1. Births and birth rates: United States, 2010–2019, and by race and Hispanic origin, United States, 2016–2019
-# Columns: Fertility Rate, All races and origins[years 2016-2018]
-
-# Hamilton BE, Martin JA, Osterman MJK. Births: Provisional data for 2020. Vital Statistics
-# Rapid Release; no 12. Hyattsville, MD: National Center for Health Statistics. May
-# 2021. DOI: https://doi.org/10.15620/cdc:104993.
-# Data can be found on page 6,
-# Table 1. Births and birth rates, by age of mother: United States, final 2019 and provisional 2020
-# Columns: 2020 Rate, 2019 Rate, Age of mother[All ages]
-
-# I crosschecked the data from these 3 reports with the data in fig_1.csv provided by 
-# the authors of the paper 
-# Kearney, Melissa S., Phillip B. Levine, and Luke Pardue. 2022. 
-# "The Puzzle of Falling US Birth Rates since the Great Recession." 
-# Journal of Economic Perspectives, 36 (1): 151-76. DOI: 10.1257/jep.36.1.151 
-# and it is the same.
+library(ggplot2)
+library(reshape2)
+library(ggplot2)
+library(readr)
+library(dplyr)
 
 
[…]

### 2024-02-15 replication/scripts/fig2_replication.R

@@ -1,96 +1,25 @@
-#### Preamble ####
-# Purpose: Figure 2 replication
-# Author: Kimlin Chin
-# Date: 22 February 2022
-# Contact: kimlin.chin@mail.utoronto.ca
-# License: MIT
 
-## Data Sources:
+library(ggplot2)
+library(reshape2)
 
-# Martin JA, Hamilton BE, Osterman MJK, et al. Births: Final data for 2015. 
-# National vital statistics report; vol 66, no 1. Hyattsville, MD: National Center for Health Statistics. 
-# 2017. https://www.cdc.gov/nchs/data/nvsr/nvsr66/nvsr66_01.pdf.
-# Data can be found on page 35,
-# Table 9. Birth rates, by live-birth order and race and Hispanic origin of mother: United States, 1980–2015
-# Columns: Fertility Rate, All races[years 1980-2015]
+library(ggplot2)
+library(reshape2)
 
-# Martin JA, Hamilton BE, Osterman MJK, Driscoll AK. Births: Final Data for 2019. National
-# Vital Statistics Reports; vol 70 no 2. Hyattsville, MD: National Center for Health
-# Statistics. 2021. DOI: https://dx.doi.org/10.15620/cdc:100472.
-# Data can be found on page 13,
-# Table 1. Births and birth rates: United States, 2010–2019, and by race and Hispanic origin, United States, 2016–2019
-# Columns: Fertility Rate, All races and origins[years 2016-2018]
+data <- read.csv("/cloud/project/inputs/data/table2.csv")
 
-# Hamilton BE, Martin JA, Osterman MJK. Births: Provisional data for 2020. Vital Statistics
-# Rapid Release; no 12. Hyattsville, MD: National Center for Health Statistics. May
-# 2021. DOI: https://doi.org/10.15620/cdc:104993.
-# Data can be found on page 6,
-# Table 1. Births and birth rates, by age of mother: United States, final 2019 and provisional 2020
-# Columns: 2020 Rate, 2019 Rate, Age of mother[All ages]
+data <- data[-1,]
+data <- data[, c(1, 3, 5)]
 
-# I crosschecked the data from these 3 reports with the data in fig_1.csv provided by 
-# the authors of the paper 
-# Kearney, Melissa S., Phillip B. Levine, and Luke Pardue. 2022. 
[…]

### 2024-02-15 replication/scripts/fig3_replication.R

@@ -1,0 +1,39 @@
+
+data <- read.csv("/cloud/project/inputs/data/osf-past-normality-regret-replication-exp1-data.csv", header = TRUE, stringsAsFactors = FALSE, fileEncoding = "UTF-8-BOM")
+data$Sc1_regret <- factor(data$Sc1_regret, levels = c(1,2), ordered = TRUE, labels = c("Exception Jones", "Routine Smith"))
+data$sc1_socnorms1 <- factor(data$sc1_socnorms1, levels = c(1,2), ordered = TRUE, labels = c("Exception Jones", "Routine Smith"))
+data$sc1_socnorms2 <- factor(data$sc1_socnorms2, levels = c(1,2), ordered = TRUE, labels = c("Exception Jones", "Routine Smith"))
+data$sc1_combinednorms <- factor(data$sc1_combinednorms, levels = c(1,2), ordered = TRUE, labels = c("Exception Jones", "Routine Smith"))
+
+data$Sc3condition1 <- NA
+data$compensationagg1 <- NA
+data$regretagg1 <- NA
+
+data <- transform(data, Sc3condition1 = ifelse(!is.na(Sc3_C1_text), 1, ifelse(!is.na(Sc3_C2_text) | !is.na(Sc3_C3_text), 2, NA)),
+                  compensationagg1 = ifelse(!is.na(Sc3_C1_text), sc3_c1_compensation, ifelse(!is.na(Sc3_C2_text), sc3_c2_compensation, ifelse(!is.na(Sc3_C3_text), sc3_c3_compensation, NA))),
+                  regretagg1 = ifelse(!is.na(Sc3_C1_text), sc3_c1_regret, ifelse(!is.na(Sc3_C2_text), sc3_c2_regret, ifelse(!is.na(Sc3_C3_text), sc3_c3_regret, NA))))
+
+data$Sc3conditionl1 <- factor(data$Sc3condition1, levels = c(1, 2), labels = c("Routine", "Exception"))
+
+data$compensationaggrecoded1 <- data$compensationagg1 - 1
+data$regretaggrecoded1 <- data$regretagg1 + 1
+
+fig3 <- ggplot(data, aes(x = Sc3conditionl1, y = compensationaggrecoded1)) + 
+  geom_violin(trim = FALSE) + 
+  geom_jitter(shape = 16, position = position_jitter(height = 0.1, width = 0.2)) + 
+  theme_apa() +
+  labs(x = "Condition", y = "Compensation") +
[…]
